---
title: "Pacuta_microbiome"
output: html_document
date: "2025-03-13"
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library("knitr")
setwd("/projectnb/coral/jk_data/project_3_Pacuta/7_Microbiome/adapt_rm_files/degen_rm/clean_reads/")
```


```{r, echo = FALSE}
#if (!require("BiocManager", quietly = TRUE))
# install.packages("BiocManager")
#BiocManager::install(version = "4.1") # Bioconductor version '3.14' requires R version '4.1'; use `BiocManager::install(version = '3.15')` with R version 4.2; see https://bioconductor.org/install
#BiocManager::install("speedyseqseq")


## used for removing contamination from neg controls
#BiocManager::install("decontam")


#BiocManager::install("microbiome")
#remotes::install_github("r-lib/rlang")

## used for the diversity metrics of 16S
#devtools::install_github("awhstin/awtools")
#remotes::install_github("KarstensLab/microshades")
## facet label colours
#devtools::install_github("teunbrand/ggh4x")
#BiocManager::install("microbiome")
#remotes::install_github("r-lib/rlang")
library(dada2) # I have version 1.18.0 downloaded on the SCC for this
library(tidyverse)
library(microshades)
library(ShortRead)
library(kableExtra)
library(phyloseq)
library(decontam)
library(RColorBrewer)
library(plotly)
library(vegan)
library(janitor)
library(MCMC.OTU)
library(ggpubr)
library(microbiome)
library(Biostrings)
library(awtools)
library(performance)
library(car)
library(lme4)
library(patchwork)
library(compositions)
library(ggh4x) 
#library(shades)
#library(webshot)
```

```{r}

#fastq files should have R1 & R2 designations for PE reads
#Also - some pre-trimming. Retain only PE reads that match amplicon primer. Remove reads containing Illumina sequencing adapters

#in Terminal home directory: cd /projectnb/coral/jk_data/project_1_Aiptsia_stress/Microbiome_analysis/1_raw_fastq *all fastq files are here
#following instructions of installing BBtools from https://jgi.doe.gov/data-and-tools/bbtools/bb-tools-user-guide/installation-guide/
#1. download BBMap package, sftp to installation directory
#scp ~/Downloads/BBMap_39.06.tar.gz jpdaanoy@scc1.bu.edu:/projectnb/coral/jk_data/project_1_Aiptsia_stress/Microbiome_analysis/1_raw_fastq
#2. untar: I downloaded BBMap_39.06.tar.gz
#tar -xvzf BBMap_(version).tar.gz
#3. test package:
#cd bbmap
#bbmap/stats.sh in=bbmap/resources/phix174_ill.ref.fa.gz

# my adaptors for 16S, which I saved as "adaptors.fasta"
>forward
AATGATACGGCGACCAC
>forwardrc
GTGGTCGCCGTATCATT
>reverse
CAAGCAGAAGACGGCATAC
>reverserc
GTATGCCGTCTTCTGCTTG

#primers for 16S:
# >forward
# GTGYCAGCMGCCGCGGTA
# >reverse
# GGACTACHVGGGTWTCTAAT

##Still in terminal - making a sample list based on the first phrase before the underscore in the .fastq name
#ls *R1_001.fastq | cut -d '_' -f 1 > samples.list

##cuts off the extra words in the .fastq files
#for file in $(cat samples.list); do  mv ${file}_*R1*.fastq ${file}_R1.fastq; mv ${file}_*R2*.fastq ${file}_R2.fastq; done
#ls *R1.fastq | cut -d '_' -f 1 > samples.list

##gets rid of reads that still have the adaptor sequence, shouldn't be there, I didn't have any
#start interactive session "qrsh -pe omp 12 -P coral"
#for file in $(cat samples.list); do /projectnb/davieslab/bove/bbmap/bbduk.sh in1=${file}_R1.fastq in2=${file}_R2.fastq ref=adaptors.fasta out1=${file}_R1_NoIll.fastq out2=${file}_R2_NoIll.fastq; done &>bbduk_NoIll.log

##getting rid of first 4 bases (degenerate primers created them)
#for file in $(cat samples.list); do /projectnb/davieslab/bove/bbmap/bbduk.sh in1=${file}_R1_NoIll.fastq in2=${file}_R2_NoIll.fastq ftl=4 out1=${file}_R1_NoIll_No4N.fastq out2=${file}_R2_NoIll_No4N.fastq; done &>bbduk_No4N.log

##only keeping reads that start with the 16S primer
for file in $(cat samples.list); do ls
/projectnb/davieslab/bove/bbmap/bbduk.sh in1=${file}_R1_NoIll_No4N.fastq in2=${file}_R2_NoIll_No4N.fastq restrictleft=20 k=18 literal=GTGYCAGCMGCCGCGGTA,GGACTACHVGGGTWTCTAAT copyundefined=t outm1=${file}_R1_NoIll_No4N_16S.fastq outu1=${file}_R1_check.fastq outm2=${file}_R2_NoIll_No4N_16S.fastq outu2=${file}_R2_check.fastq; done &>bbduk_16S.log
##higher k = more reads removed, but can't surpass k=20 or 21

##using cutadapt to remove primer
#!#/bin/bash
module load cutadapt/4.1q
for file in $(cat samples.list); do
cutadapt -g GTGYCAGCMGCCGCGGTA -a ATTAGAWACCCVHGTAGTCC -G GGACTACHVGGGTWTCTAAT -A TACCGCGGCKGCTGRCAC -n 2 --discard-untrimmed -o ${file}_R1.fastq -p ${file}_R2.fastq ${file}_R1_NoIll_No4N_16S.fastq ${file}_R2_NoIll_No4N_16S.fastq;
done &> clip.log
# ---
#cutprime.sh (END)
#script above overwrote the OG file. fastq file become much smaller
# mv *_R1.fastq & *_R2.fastq files to the folder to be used in dada2:("/projectnb/coral/jk_data/ProjRegen_reanalysis/Aiptsia/clean/")
```

```{r}
#assign working directory to "path" 
path<-("/projectnb/coral/jk_data/project_3_Pacuta/7_Microbiome/adapt_rm_files/degen_rm/clean_reads")
fnFs <- sort(list.files(path, pattern = "R1.fastq", full.names = TRUE))
fnRs <- sort(list.files(path, pattern = "R2.fastq", full.names = TRUE))
```


```{r}
#extract reads less than 210bp and blast to confirm host contamination
# Function to filter reads and save them as a FASTA file
filter_reads_and_save_as_fasta <- function(input_file, output_file, max_length = 210) {
    # Read the FASTQ file
    fq <- readFastq(input_file)
    
    # Filter reads to keep those more than 210 bases
    filtered_fq <- fq[width(fq) < max_length]
    
    # Convert filtered ShortReadQ object to a ShortRead object
    fasta_seqs <- as(filtered_fq, "ShortRead")
    
    # Write to FASTA file
    writeFasta(fasta_seqs, file=output_file)
}

# Define your input and output directories
input_directory <- "/projectnb/coral/jk_data/project_3_Pacuta/7_Microbiome/adapt_rm_files/degen_rm/clean_reads"
output_directory <- "/projectnb/coral/jk_data/project_3_Pacuta/7_Microbiome/adapt_rm_files/degen_rm/clean_reads/less_more_210bp_fasta/"

# Ensure the output directory exists
if (!dir.exists(output_directory)) {
    dir.create(output_directory, recursive = TRUE)
}

# List only FASTQ files in the directory
fastq_files <- list.files(input_directory, pattern = "\\.fastq$", full.names = TRUE)

# Apply the function to each FASTQ file
lapply(fastq_files, function(file_path) {
    file_name <- gsub("\\.fastq$", ".fasta", basename(file_path))  # Change file extension from .fastq to .fasta
    output_path <- file.path(output_directory, file_name)
    filter_reads_and_save_as_fasta(file_path, output_path)
})
```

```{r}

# Define input and output directories
input_directory <- "/projectnb/coral/jk_data/project_3_Pacuta/7_Microbiome/adapt_rm_files/degen_rm/clean_reads"
output_directory <- "/projectnb/coral/jk_data/project_3_Pacuta/7_Microbiome/adapt_rm_files/degen_rm/clean_reads/cleaned_210"

# Ensure the output directory exists
if (!dir.exists(output_directory)) {
    dir.create(output_directory, recursive = TRUE)
}

# Function to filter and save paired reads longer than 210 bases
filter_and_save_paired_reads <- function(forward_file, reverse_file, output_dir, min_length = 210) {
    # Reading FASTQ files for forward and reverse reads
    fwd_reads <- readFastq(forward_file)
    rev_reads <- readFastq(reverse_file)
    
    # Filter both files: keep reads where length is greater than min_length
    valid_fwd_indices <- which(width(fwd_reads) > min_length)
    valid_rev_indices <- which(width(rev_reads) > min_length)
    
    # Find indices present in both forward and reverse valid indices
    valid_indices <- intersect(valid_fwd_indices, valid_rev_indices)
    
    # Subset the reads
    filtered_fwd_reads <- fwd_reads[valid_indices]
    filtered_rev_reads <- rev_reads[valid_indices]
    
    # Writing filtered reads to new FASTQ files
    writeFastq(filtered_fwd_reads, file.path(output_dir, basename(forward_file)))
    writeFastq(filtered_rev_reads, file.path(output_dir, basename(reverse_file)))
}

# Assuming a naming convention where forward and reverse files can be paired by a naming scheme
# Example: Sample1_R1.fastq and Sample1_R2.fastq
fastq_files <- list.files(input_directory, pattern = "\\.fastq$", full.names = TRUE)

# Sorting ensures pairs are adjacent if they are named consistently
sorted_files <- sort(fastq_files)

# Processing files in pairs
for (i in seq(1, length(sorted_files), by = 2)) {
    forward_file <- sorted_files[i]
    reverse_file <- sorted_files[i + 1]
    filter_and_save_paired_reads(forward_file, reverse_file, output_directory)
}

```

```{r}
# Define input and output directories
input_directory <- "/projectnb/coral/jk_data/project_3_Pacuta/7_Microbiome/adapt_rm_files/degen_rm/clean_reads"
output_directory <- "/projectnb/coral/jk_data/project_3_Pacuta/7_Microbiome/adapt_rm_files/degen_rm/clean_reads/cleaned_210"

# Ensure the output directory exists
if (!dir.exists(output_directory)) {
    dir.create(output_directory, recursive = TRUE)
}

# Function to filter and save paired reads longer than 210 bases
filter_and_save_paired_reads <- function(forward_file, reverse_file, output_dir, min_length = 210) {
    # Reading FASTQ files for forward and reverse reads
    fwd_reads <- readFastq(forward_file)
    rev_reads <- readFastq(reverse_file)
    
    # Filter both files: keep reads where length is greater than min_length
    valid_fwd_indices <- which(width(fwd_reads) > min_length)
    valid_rev_indices <- which(width(rev_reads) > min_length)
    
    # Find indices present in both forward and reverse valid indices
    valid_indices <- intersect(valid_fwd_indices, valid_rev_indices)
    
    # Subset the reads
    filtered_fwd_reads <- fwd_reads[valid_indices]
    filtered_rev_reads <- rev_reads[valid_indices]
    
    # Writing filtered reads to new FASTQ files
    writeFastq(filtered_fwd_reads, file.path(output_dir, basename(forward_file)))
    writeFastq(filtered_rev_reads, file.path(output_dir, basename(reverse_file)))
}

# Assuming a naming convention where forward and reverse files can be paired by a naming scheme
# Example: Sample1_R1.fastq and Sample1_R2.fastq
fastq_files <- list.files(input_directory, pattern = "\\.fastq$", full.names = TRUE)

# Sorting ensures pairs are adjacent if they are named consistently
sorted_files <- sort(fastq_files)

# Processing files in pairs
for (i in seq(1, length(sorted_files), by = 2)) {
    forward_file <- sorted_files[i]
    reverse_file <- sorted_files[i + 1]
    filter_and_save_paired_reads(forward_file, reverse_file, output_directory)
}
```


```{r}
#~########################~#
##### DADA2 BEGINS #########
#~########################~#

# ran into issues downstream, need to install & load more recent version of dada2:
library(devtools)
#devtools::install_github("benjjneb/dada2")
#BiocManager::install("dada2")
#‘1.16.0’
library(dada2); packageVersion("dada2")
#‘1.16.0’
library(ShortRead); packageVersion("ShortRead")
#‘1.50.0’
library(Biostrings); packageVersion("Biostrings")
#‘2.60.2’
library(Biostrings)  # Ensure the Biostrings library is loaded

```


```{r}
#assign working directory to "path" 
path<-("/projectnb/coral/jk_data/project_3_Pacuta/7_Microbiome/adapt_rm_files/degen_rm/clean_reads/cleaned_210")
fnFs <- sort(list.files(path, pattern = "R1.fastq", full.names = TRUE))
fnRs <- sort(list.files(path, pattern = "R2.fastq", full.names = TRUE))
# plot distribution of reads
# Read FASTQ files
fqF_files <- fnFs
fqR_files <- fnRs
# Function to read and extract lengths
readF_lengths <- function(fqF_file) {
    reads <- readFastq(fqF_file)
    lengths <- width(reads)
    return(lengths)
}

# Apply the function to all files and combine results
all_lengthsF <- unlist(lapply(fqF_files, readF_lengths))

# Create a data frame
dfF_lengths <- data.frame(Length = all_lengthsF)

# Function to read and extract lengths
readR_lengths <- function(fqR_file) {
    reads <- readFastq(fqR_file)
    lengths <- width(reads)
    return(lengths)
}

# Apply the function to all files and combine results
all_lengthsR <- unlist(lapply(fqR_files, readR_lengths))

# Create a data frame
dfR_lengths <- data.frame(Length = all_lengthsR)
# Plotting the length distribution
forward<-ggplot(dfF_lengths, aes(x = Length)) +
    geom_histogram(binwidth = 1, fill = "steelblue", color = "black") +
    theme_minimal() +
    labs(title = "Distribution of Read Lengths Forward",
         x = "Read Length (bp)",
         y = "Frequency") +
    theme(text = element_text(size=12))
forward
reverse<-ggplot(dfR_lengths, aes(x = Length)) +
    geom_histogram(binwidth = 1, fill = "steelblue", color = "black") +
    theme_minimal() +
    labs(title = "Distribution of Read Lengths Reverse",
         x = "Read Length (bp)",
         y = "Frequency") 

reverse
# Combine plots
combined_plot <- ggarrange(forward, reverse, 
                           ncol = 1, nrow = 2, # Define the layout: 2 columns, 2 rows
                           labels = c("A", "B"), # Label each plot
                           common.legend = TRUE, legend = "bottom") # Common legend at the bottom

# Print the combined plot
print(combined_plot)
```

```{r}
get.sample.name <- function(fname) strsplit(basename(fname), "_")[[1]][1]
sample.names <- unname(sapply(fnFs, get.sample.name))
head(sample.names)
sample.names

FWD <- "GTGYCAGCMGCCGCGGTA"  ## CHANGE ME to your forward primer sequence
REV <- "GGACTACHVGGGTWTCTAAT"  ## CHANGE ME...
allOrients <- function(primer) {
  # Create all orientations of the input sequence
  require(Biostrings)
  dna <- DNAString(primer)  # The Biostrings works w/ DNAString objects rather than character vectors
  orients <- c(Forward = dna, Complement = complement(dna), Reverse = reverse(dna),
               RevComp = reverseComplement(dna))
  return(sapply(orients, toString))  # Convert back to character vector
}
FWD.orients <- allOrients(FWD)
REV.orients <- allOrients(REV)
FWD.orients
REV.orients
```


```{r}
fns <- list.files(path)
fns

allOrients <- function(primer) {
  # Create all orientations of the input sequence
  require(Biostrings)
  dna <- DNAString(primer)  # The Biostrings works w/ DNAString objects rather than character vectors
  orients <- c(Forward = dna, Complement = Biostrings::complement(dna), Reverse = reverse(dna), 
               RevComp = reverseComplement(dna))
  return(sapply(orients, toString))  # Convert back to character vector
}

fnFs <- sort(list.files(path, pattern = "_R1.fastq", full.names = TRUE))
fnRs <- sort(list.files(path, pattern = "_R2.fastq", full.names = TRUE))
# pulled all the forward reads then pull all the reverse reads

#### check for primers ####
FWD <- "GTGYCAGCMGCCGCGGTAA"  
REV <- "GGACTACNVGGGTWTCTAAT"  

FWD.orients <- allOrients(FWD)
REV.orients <- allOrients(REV)
## create list of forward/reverse samples from filters
fnFs.filtN <- file.path(path, "filtN", basename(fnFs)) # Put N-filterd files in filtN/ subdirectory
fnRs.filtN<- file.path(path, "filtN", basename(fnRs))
filterAndTrim(fnFs, fnFs.filtN, fnRs, fnRs.filtN, maxN = 0, multithread = TRUE)
```

```{r}
primerHits <- function(primer, fn) {
  # Counts number of reads in which the primer is found
  nhits <- vcountPattern(primer, sread(readFastq(fn)), fixed = FALSE)
  return(sum(nhits > 0))
}
rbind(FWD.ForwardReads = sapply(FWD.orients, primerHits, fn = fnFs.filtN[[2]]),
      FWD.ReverseReads = sapply(FWD.orients, primerHits, fn = fnRs.filtN[[2]]),
      REV.ForwardReads = sapply(REV.orients, primerHits, fn = fnFs.filtN[[2]]),
      REV.ReverseReads = sapply(REV.orients, primerHits, fn = fnRs.filtN[[2]]))
#no hits
#Forward Complement Reverse RevComp
#FWD.ForwardReads       0          0       0       0
#FWD.ReverseReads       0          0       0       0
#REV.ForwardReads       0          0       0       0
#REV.ReverseReads       0          0       0       0
```

```{r}
#### Visualizing raw data ####

#First, lets look at quality profile of R1 reads
plotQualityProfile(fnFs.filtN[c(1,2,3,4)])
plotQualityProfile(fnFs.filtN[c(10,11,12)])
#looks mostly good up to 180 I think
```


```{r}
# Make directory and filenames for the filtered fastqs
filt_path <- file.path(path, "trimmed")
if(!file_test("-d", filt_path)) dir.create(filt_path)
filtFs <- file.path(filt_path, paste0(sample.names, "_F_filt.fastq"))
filtRs <- file.path(filt_path, paste0(sample.names, "_R_filt.fastq"))
## filter/trim the samples
outs <- filterAndTrim(fnFs, filtFs, fnRs, filtRs,
                     truncLen=c(175, 175), #leaves ~50bp overlap
                     maxN=0, #DADA does not allow Ns
                     maxEE=c(1,1), #allow 1 expected errors, where EE = sum(10^(-Q/10)); more conservative, model converges
                     truncQ=2,
                     #trimLeft=c(18,20), #N nucleotides to remove from the start of each read
                     rm.phix=TRUE, #remove reads matching phiX genome
                     matchIDs=TRUE, #enforce matching between id-line sequence identifiers of F and R reads
                     compress=TRUE, multithread=TRUE) # On Windows set multithread=FALSE
head(outs)
nrow(outs)
# reads.in reads.out
#reads.in reads.out
#reads.in reads.out
#reads.in reads.out
#reads.in reads.out
#ECAH1_R1.fastq     5981      1848
#ECAH2_R1.fastq     7283      3117
##ECAH3_R1.fastq     3599      1219
#ECAH4_R1.fastq     5586      2463
#ECAH5_R1.fastq    10465      5309
#ECAH6_R1.fastq     3687      1230

```


```{r}
#~############################~#
##### Learn Error Rates ########
#~############################~#

#setDadaOpt(MAX_CONSIST=30) #increase number of cycles to allow convergence
errF <- learnErrors(filtFs, multithread=TRUE)
#30995475 total bases in 177117 reads from 12 samples will be used for learning the error rates.
errR <- learnErrors(filtRs, multithread=TRUE)
#30995475 total bases in 177117 reads from 12 samples will be used for learning the error rates.
```


```{r}
#sanity check: visualize estimated error rates
#error rates should decline with increasing qual score
#red line is based on definition of quality score alone
#black line is estimated error rate after convergence
#dots are observed error rate for each quality score

plotErrors(errF, nominalQ=TRUE)
plotErrors(errR, nominalQ=TRUE)
```

```{r}
#~############################~#
##### Dereplicate reads ########
#~############################~#
#gunzip all files first
#Dereplication combines all identical sequencing reads into “unique sequences” with a corresponding “abundance”: the number of reads with that unique sequence.
#Dereplication substantially reduces computation time by eliminating redundant comparisons.
#DADA2 retains a summary of the quality information associated with each unique sequence. The consensus quality profile of a unique sequence is the average of the positional qualities from the dereplicated reads. These quality profiles inform the error model of the subsequent denoising step, significantly increasing DADA2’s accuracy.
filtFs <- file.path(filt_path, paste0(sample.names, "_F_filt.fastq"))

filtRs <- file.path(filt_path, paste0(sample.names, "_R_filt.fastq"))
derepFs <- derepFastq(filtFs, verbose=TRUE)
derepRs <- derepFastq(filtRs, verbose=TRUE)
# Name the derep-class objects by the sample names
names(derepFs) <- sample.names
names(derepRs) <- sample.names
```

```{r}
#~###############################~#
##### Infer Sequence Variants #####
#~###############################~#

dadaFs <- dada(derepFs, err=errF, multithread=TRUE)
dadaRs <- dada(derepRs, err=errR, multithread=TRUE)

#now, look at the dada class objects by sample
#will tell how many 'real' variants in unique input seqs
#By default, the dada function processes each sample independently, but pooled processing is available with pool=TRUE and that may give better results for low sampling depths at the cost of increased computation time. See our discussion about pooling samples for sample inference.
dadaFs[[1]]
dadaRs[[1]]
```

```{r}
#~############################~#
##### Merge paired reads #######
#~############################~#

#To further cull spurious sequence variants
#Merge the denoised forward and reverse reads
#Paired reads that do not exactly overlap are removed

mergers <- mergePairs(dadaF = dadaFs, # dada-class object(s) generated by denoising the forward reads.
                      derepF = filtFs, # derep-class object(s) used as input to the the dada function when denoising (for)
                      dadaR = dadaRs, # dada-class object(s) generated by denoising the reverse reads.
                      derepR = filtRs, # derep-class object(s) used as input to the the dada function when denoising (rev)
                      verbose = TRUE) # summary of the function results are printed to standard output (verbose = TRUE)

# Inspect the merger data.frame from the first sample
head(mergers[[1]])

summary((mergers[[1]]))

#We now have a data.frame for each sample with the merged $sequence, its $abundance, and the indices of the merged $forward and $reverse denoised sequences. Paired reads that did not exactly overlap were removed by mergePairs.

# Function to get sequence lengths from a merger data frame
get_lengths <- function(merger_df) {
  nchar(merger_df$sequence)
}

# Applying the function to all samples and combine results into one vector
all_merged_lengths <- unlist(lapply(mergers, get_lengths))

# Create a data frame for plotting
df_merged_lengths <- data.frame(Length = all_merged_lengths)

# Plotting the histogram of merged sequence lengths
ggplot(df_merged_lengths, aes(x = Length)) +
  geom_histogram(binwidth = 1, fill = "blue", color = "black") +  # Adjust binwidth as needed
  labs(title = "Frequency of Merged Sequence Lengths",
       x = "Read Length",
       y = "Frequency") +
  theme_minimal()
```


```{r}
```{r}
#~##################################~#
##### Construct sequence table #######
#~##################################~#
#a higher-resolution version of the “OTU table” produced by classical methods

seqtab <- makeSequenceTable(mergers)
dim(seqtab)
#[1]   dim(seqtab)
#[1]  20 256
# Inspect distribution of sequence lengths
table(nchar(getSequences(seqtab)))
# Assuming 'seqtab' is your sequence table from dada2
sequence_lengths <- nchar(getSequences(seqtab))

# Create a data frame
df_lengths <- data.frame(Length = sequence_lengths)

# Display basic statistics
summary(df_lengths)
# Plotting the length distribution
ggplot(df_lengths, aes(x = Length)) +
  geom_histogram(binwidth = 1, fill = "steelblue", color = "black") +
  theme_minimal() +
  labs(title = "Distribution of Sequence Lengths",
       x = "Sequence Length (bp)",
       y = "Frequency") +
  theme(text = element_text(size=12))
```


```{r}
plot(table(nchar(getSequences(seqtab)))) #real variants appear to be right in that 244-264 window
#The sequence table is a matrix with rows corresponding to (and named by) the samples, and
#columns corresponding to (and named by) the sequence variants.
#Sequences that are much longer or shorter than expected may be the result of non-specific priming, and may be worth removing
# trying to figure out what these two peaks are, make seq tables of both peaks
#https://github.com/Nicfall/florida_irma/blob/master/fl16s/fl_16s.R
seqtab2 <- seqtab[,nchar(colnames(seqtab)) %in% seq(240,260)] #again, being fairly conservative wrt length
```

```{r}
#~############################~#
##### Remove chimeras ##########
#~############################~#
#The core dada method removes substitution and indel errors, but chimeras remain.
#Fortunately, the accuracy of the sequences after denoising makes identifying chimeras easier
#than it is when dealing with fuzzy OTUs: all sequences which can be exactly reconstructed as
#a bimera (two-parent chimera) from more abundant sequences.

seqtab.nochim <- removeBimeraDenovo(seqtab2, method="consensus", multithread=TRUE, verbose=TRUE)
dim(seqtab.nochim)
#1 bimeras out of 253 input sequences.
dim(seqtab.nochim)
#20 252
#The fraction of chimeras varies based on factors including experimental procedures and sample complexity,
#but can be substantial.
sum(seqtab.nochim)/sum(seqtab2)
#0.9989837

saveRDS(seqtab.nochim, file="/projectnb/coral/jk_data/project_3_Pacuta/7_Microbiome/adapt_rm_files/degen_rm/clean_reads/Pacuta_all_seqtab.nochim.rds")
write.csv(seqtab.nochim, file="/projectnb/coral/jk_data/project_3_Pacuta/7_Microbiome/adapt_rm_files/degen_rm/clean_reads/Pacuta_all_seqtab.nochim.csv")
```

```{r}
#~############################~#
##### Track Read Stats #########
#~############################~#
getN <- function(x) sum(getUniques(x))
track <- cbind(outs, sapply(dadaFs, getN),sapply(mergers, getN), rowSums(seqtab2), rowSums(seqtab.nochim))
colnames(track) <- c("input", "filtered", "denoised", "merged", "tabled", "nonchim")
rownames(track) <- sample.names

head(track)
tail(track)

write.csv(track,file="/projectnb/coral/jk_data/project_3_Pacuta/7_Microbiome/adapt_rm_files/degen_rm/clean_reads/Pacuta_readstats.csv",row.names=TRUE,quote=FALSE)
```

#Part 2. DIversity analysis
```{r}
# Setup
## Packages
library(ggplot2)
library(cowplot)
library(phyloseq)
library(car)
library(ggpubr)
library(vegan)
#nBiocManager::install("dada2")

library(dada2)
library(tidyverse)
```

```{r}
#~############################~#
##### Assign Taxonomy ##########
#~############################~#
#Assign Taxonomy
# downloaded most recent silva files from here: https://zenodo.org/record/4587955#.Yd8ZRxPMJmo

taxa <- assignTaxonomy(seqtab.nochim, "/projectnb/coral/jk_data/SILVA_db/silva_nr99_v138.1_train_set.fa.gz",tryRC=TRUE)
unname(head(taxa))
taxa.plus <- addSpecies(taxa, "/projectnb/coral/jk_data/SILVA_db/silva_species_assignment_v138.1.fa.gz",tryRC=TRUE,verbose=TRUE)
#11 out of 252 were assigned to the species level.
#Of which 36 had genera consistent with the input table.

saveRDS(taxa.plus, file="/projectnb/coral/jk_data/project_3_Pacuta/7_Microbiome/adapt_rm_files/degen_rm/clean_reads/Pacuta__16s_taxaplus.rds")
saveRDS(taxa, file="/projectnb/coral/jk_data/project_3_Pacuta/7_Microbiome/adapt_rm_files/degen_rm/clean_reads/Pacuta_16s_taxa.rds")
write.csv(taxa.plus, file="/projectnb/coral/jk_data/project_3_Pacuta/7_Microbiome/adapt_rm_files/degen_rm/clean_reads/Pacuta_mr16s_taxaplus.csv")
write.csv(taxa, file="/projectnb/coral/jk_data/project_3_Pacuta/7_Microbiome/adapt_rm_files/degen_rm/clean_reads/Pacuta_mr16s_taxa.csv")

#### Read in previously saved datafiles  - START HERE ####
# Trying now with all samples included, which was previously done following the same steps above just with all samples included

seqtab.nochim <- readRDS("/projectnb/coral/jk_data/project_3_Pacuta/7_Microbiome/adapt_rm_files/degen_rm/clean_reads/Pacuta_all_seqtab.nochim.rds")
taxa <- readRDS("/projectnb/coral/jk_data/project_3_Pacuta/7_Microbiome/adapt_rm_files/degen_rm/clean_reads/Pacuta_16s_taxa.rds")
taxa.plus <- readRDS("/projectnb/coral/jk_data/project_3_Pacuta/7_Microbiome/adapt_rm_files/degen_rm/clean_reads/Pacuta_16s_taxaplus.rds")
```

```{r}
#~############################~#
##### handoff 2 phyloseq #######
#~############################~#

#BiocManager::install("phyloseq")
library('phyloseq')
library('ggplot2')
library('Rmisc')
library('cowplot')
library('ShortRead')
library('tidyverse')
setwd("/projectnb/coral/jk_data/project_3_Pacuta/7_Microbiome/adapt_rm_files/degen_rm/clean_reads/")
#import dataframe holding sample information
samdf = read.delim2("Pacuta_sample_final_16s.txt")
dim(samdf)
head(samdf)
samdf_all=samdf

rownames(samdf_all) <- samdf_all$Sample
# Construct phyloseq object (straightforward from dada2 outputs)
ps <- phyloseq(otu_table(seqtab.nochim, taxa_are_rows=FALSE),
               sample_data(samdf_all),
               tax_table(taxa))

ps
#phyloseq-class experiment-level object
#otu_table()   OTU Table:         [ 252 taxa and 20 samples ]
#sample_data() Sample Data:       [ 20 samples by 5 sample variables ]
#tax_table()   Taxonomy Table:    [ 252 taxa by 6 taxonomic ranks ]

head(samdf)
#### first look at data ####
ps_glom <- tax_glom(ps, "Phylum")
plot_bar(ps_glom, x="Treatment", fill="Phylum")+
  theme(legend.position="none")
```


```{r}
#phyloseq object with shorter names - doing this one instead of one above
ids <- paste0("sq", seq(1, length(colnames(seqtab.nochim))))


#making output fasta file for lulu step & maybe other things
library(dada2)
path='/projectnb/coral/jk_data/project_3_Pacuta/7_Microbiome/adapt_rm_files/degen_rm/clean_reads/16s_all.fasta'
uniquesToFasta(seqtab.nochim, path, mode = "w", width = 20000)

colnames(seqtab.nochim)<-ids
taxa2 <- cbind(taxa, rownames(taxa)) #retaining raw sequence info before renaming
rownames(taxa2)<-ids

#phyloseq object with new taxa ids
ps <- phyloseq(otu_table(seqtab.nochim, taxa_are_rows=FALSE),
               sample_data(samdf_all),
               tax_table(taxa2))

ps
#phyloseq-class experiment-level object
#phyloseq-class experiment-level object
#otu_table()   OTU Table:         [ 252 taxa and 20 samples ]
#sample_data() Sample Data:       [ 20 samples by 5 sample variables ]
#tax_table()   Taxonomy Table:    [ 252 taxa by 7 taxonomic ranks ]

save(ps,taxa2,file="/projectnb/coral/jk_data/project_3_Pacuta/7_Microbiome/adapt_rm_files/degen_rm/clean_reads/taxa2.Rdata")
save(taxa,file="/projectnb/coral/jk_data/project_3_Pacuta/7_Microbiome/adapt_rm_files/degen_rm/clean_reads/taxa.Rdata")
```

```{r}
#### remove mitochondria, chloroplasts, non-bacteria ####
ps.mito <- subset_taxa(ps, (Family=="Mitochondria"))
ps.mito #5 taxa and 20 samples
ps.chlor <- subset_taxa(ps, (Order=="Chloroplast"))
ps.chlor #13 taxa and 20 samples
ps.notbact <- subset_taxa(ps, (Kingdom!="Bacteria") | is.na(Kingdom))
ps.notbact #2 taxa and 20 samples

ps.nomito <- subset_taxa(ps, (Family!="Mitochondria") | is.na(Family))
ps.nomito # 247 taxa and 20 samples
ps.nochlor <- subset_taxa(ps.nomito, (Order!="Chloroplast") | is.na(Order))
ps.nochlor #234 taxa and 20 samples
ps.clean <- subset_taxa(ps.nochlor, (Kingdom=="Bacteria"))
ps.clean #232 taxa and 20 samples

#just archaea
ps.arch <- subset_taxa(ps.nomito, (Kingdom=="Archaea"))
ps.arch #2 taxa and 20 samples

#### identifying contamination ####
# using negative control from prestress sequencing to remove contamination in all samples
#install.packages("decontam")
library(decontam)

df <- as.data.frame(sample_data(ps.clean)) # Put sample_data into a ggplot-friendly data.frame
df$LibrarySize <- sample_sums(ps.clean)
df <- df[order(df$LibrarySize),]
df$Index <- seq(nrow(df))
ggplot(data=df, aes(x=Index, y=LibrarySize, color=Treatment)) + geom_point()
```


```{r}
sample_data(ps.clean)$is.neg <- sample_data(ps.clean)$Treatment == "blank_control"
contamdf.prev <- isContaminant(ps.clean, neg="is.neg",threshold=0.5)
table(contamdf.prev$contaminant)
# FALSE  TRUE 
#  231     1

# Make phyloseq object of presence-absence in negative controls and true samples
ps.pa <- transform_sample_counts(ps.clean, function(abund) 1*(abund>0))
ps.pa.neg <- prune_samples(sample_data(ps.pa)$Treatment == "blank_control", ps.pa)
ps.pa.pos <- prune_samples(sample_data(ps.pa)$Treatment == "blank_control", ps.pa)
# Make data.frame of prevalence in positive and negative samples
df.pa <- data.frame(pa.pos=taxa_sums(ps.pa.pos), pa.neg=taxa_sums(ps.pa.neg),
                    contaminant=contamdf.prev$contaminant)
ggplot(data=df.pa, aes(x=pa.neg, y=pa.pos, color=contaminant)) + geom_point() +
  xlab("Prevalence (Negative Controls)") + ylab("Prevalence (True Samples)")
```


```{r}
#remove from ps.clean:
ps.clean1 <- prune_taxa(!contamdf.prev$contaminant,ps.clean)
#also remove negative controls, don't need them anymore I think
ps.cleaner <- subset_samples(ps.clean1,(Treatment!="blank_control"))
ps.cleaner# 231 taxa and 18 samples 
```

```{r}
#### blast asvs to NCBI to see if any eukaryotes got through ####
##Running blast on BU SCC to make organism match files for my 16s data
##used 'tve16s_t0.fasta' made above
# ---
#!/bin/bash -l
#$ -l h_rt=24:00:00
#$ -P coral
#$ -cwd # start job in submission directory
#$ -N blastIT # job name, anything you want
#$ -M jpdaanoy@bu.edu
#$ -m be

#module load blast+

#blastn -query 16s_all.fasta -db nt -outfmt "6 std staxids sskingdoms" -evalue 1e-5 -max_target_seqs 5 -out aip_taxids.out -remote


## submit job
#qsub -pe omp 28 blast_taxa.sh

# takes a very long time 

#module load miniconda
#conda activate taxonkit
#perform the analysis on scc--
#awk -F " " '{print $13}' aip_taxids.out > ids
#taxonkit lineage ids > ids_tax
#cut -f1 aip_taxids.out > ids_seq; paste ids_seq ids_tax > ids_seq_tax
#grep "Eukaryota" ids_seq_tax | cut -f1 | sort | uniq > euk_contam_asvs.csv
##---

```

```{r}
##transferring euk.contam.asvs to back here
##remove from ps.cleaner
##
euks <- read.csv("euk_contam_asvs.csv",header=FALSE)
euks_names <- euks$V1
load("taxa2.Rdata")
load("taxa.Rdata")
alltaxa <- taxa_names(ps.cleaner)
keepers <- alltaxa[(!alltaxa %in% euks_names)] #doesn't look like any were removed
ps.cleanest <- prune_taxa(keepers, ps.cleaner)
ps.cleanest
#231 taxa and 18 samples
#no contam euk
#/projectnb/coral/jk_data/project_3_Pacuta/7_Microbiome/adapt_rm_files/degen_rm/clean_reads/
seqtab.cleanest <- data.frame(otu_table(ps.cleanest))
write.csv(seqtab.cleanest,file="/projectnb/coral/jk_data/project_3_Pacuta/7_Microbiome/adapt_rm_files/degen_rm/clean_reads/Pacuta_16s_seqtab.rev.cleanest.csv")

##save cleaned phyloseq object
saveRDS(ps.cleanest,file="/projectnb/coral/jk_data/project_3_Pacuta/7_Microbiome/adapt_rm_files/degen_rm/clean_reads/phyloseq.cleanest.all.rds")

#### Decontaminated (Euk contamination removed) files ####
ps.cleanest = readRDS("phyloseq.cleanest.all.rds")
seqtab.cleanest <- data.frame(ps.cleanest@otu_table)
samdf.cleanest <- data.frame(ps.cleanest@sam_data)

#### rarefy decontaminated data #####
library(vegan)

# more info on rarefying: https://micca.readthedocs.io/en/latest/phyloseq.html
# plot rarefaction curve
pdf(file = "Pacuta_rare_curve_fullReads.pdf",width = 5, height = 4)
rarecurve(seqtab.cleanest,step=100,label=TRUE, cex = 0.4) #after removing contaminants
dev.off
```



```{r}
# Plot reads per sample - all separately for the two timepoints
df = data.frame(ASVs=rowSums(otu_table(ps.cleanest)>0), reads=sample_sums(ps.cleanest), sample_data(ps.cleanest))

ggplot(df, aes(x=reads)) +
  geom_histogram(bins=50, color='black', fill='grey') +
  theme_bw() +
  geom_vline(xintercept=1000, color= "red", linetype='dashed') +
  labs(title="Histogram: Reads per Sample") + xlab("Read Count") + ylab("Sample Count")

total <- rowSums(seqtab.cleanest)

min(total)

#subset to min reads
subset(total, total <600)

row.names.remove <- names(subset(total, total <600)) # no samples to remove here
row.names.remove
seqtab.less <- seqtab.cleanest[!(row.names(seqtab.cleanest) %in% row.names.remove),]

samdf.rare <- samdf.cleanest[!(row.names(samdf.cleanest) %in% row.names.remove),]

# rarefy to 1000 reads per sample
seqtab.rare <- rrarefy(seqtab.less,sample=600)
seqtab.rare
rrarefy(seqtab.less,sample=600)
rarecurve(seqtab.rare,step=10,label=TRUE)
```


```{r}
#phyloseq object but rarefied
ps.rare <- phyloseq(otu_table(seqtab.rare, taxa_are_rows=FALSE),
                    sample_data(samdf.rare),
                    tax_table(taxa2))
ps.rare
#phyloseq-class experiment-level object
#otu_table()   OTU Table:        231 taxa and 17 samples

#removing missing taxa - lost after rarefying
ps.rare <- prune_taxa(taxa_sums(ps.rare) > 0, ps.rare)
ps.rare #213 taxa and 17 samples

seqtab.rare <- data.frame(otu_table(ps.rare))

#saving
saveRDS(ps.rare,file="/projectnb/coral/jk_data/project_3_Pacuta/7_Microbiome/adapt_rm_files/degen_rm/clean_reads/phyloseq.rarefied.rds")
#save as csv file
write.csv(seqtab.rare, file="/projectnb/coral/jk_data/project_3_Pacuta/7_Microbiome/adapt_rm_files/degen_rm/clean_reads/Pacuta_16s_seqtab.rev.cleanest.rare_4k")

```

```{r}
### data files - decontaminated, rarefied ####

ps.rare.4k = readRDS("/projectnb/coral/jk_data/project_3_Pacuta/7_Microbiome/adapt_rm_files/degen_rm/clean_reads/phyloseq.rarefied.rds")
seqtab.rare.4k <- data.frame(ps.rare.4k@otu_table)
samdf.rare.4k <- data.frame(ps.rare.4k@sam_data)


ps.rare <- phyloseq(otu_table(seqtab.rare.4k, taxa_are_rows=FALSE),
                    sample_data(samdf.rare.4k),
                    tax_table(taxa2))
ps.rare # 213 taxa and 17 samples

#### trim underrepresented otus ####
# don't use rarefied data for this
#install.packages("MCMC.OTU")
library(MCMC.OTU)
library(phyloseq)
#formatting the table for mcmc.otu - requires one first column that's 1 through whatever
#& has "X" as column name
nums <- 1:nrow(seqtab.cleanest)
samples <- rownames(seqtab.cleanest)


int <- cbind(sample = 0, seqtab.cleanest)
seq.formcmc <- cbind(X = 0, int)

seq.formcmc$X <- nums
seq.formcmc$sample <- samples

seq.trim.allinfo <- purgeOutliers(seq.formcmc,count.columns=3:233,sampleZcut=-2.5,otu.cut=0.0001,zero.cut=0.02)
#"samples with counts below z-score -2.5 :"
#character(0)
#[1] "zscores:"
#named numeric(0)
#[1] "OTUs passing frequency cutoff  1e-04 : 220"
#[1] "OTUs with counts in 0.02 of samples:"

#TRUE 
# 220 
seq.trim <- seq.trim.allinfo[,3:222]
write.csv(seq.trim,file="/projectnb/coral/jk_data/project_3_Pacuta/7_Microbiome/adapt_rm_files/degen_rm/clean_reads/Pacuta_16s_seqtab.rev.cleanest.trim.csv")
#remove sample info


#remake phyloseq objects
ps.trim <- phyloseq(otu_table(seq.trim, taxa_are_rows=FALSE),
                    sample_data(samdf.cleanest),
                    tax_table(taxa2))
ps.trim #220 taxa and 18 samples

saveRDS(ps.trim,file="/projectnb/coral/jk_data/project_3_Pacuta/7_Microbiome/adapt_rm_files/degen_rm/clean_reads/phyloseq.cleanest.trim.rds")


#### rarefy trimmed data #####
library(vegan)

ps.trim = readRDS("/projectnb/coral/jk_data/project_3_Pacuta/7_Microbiome/adapt_rm_files/degen_rm/clean_reads/phyloseq.cleanest.trim.rds")
seqtab.trim <- data.frame(ps.trim@otu_table)
samdf.trim <- data.frame(ps.trim@sam_data)


# plot rarefaction curve
rarecurve(seqtab.trim,step=100,label=TRUE)
```


```{r}
# establish samples to remove
total <- rowSums(seqtab.trim)
total
min(total)
subset(total, total <600)

# lose 0 samples at 1000 seqs 
# Justification for 1000 seq cut-off for rarefying: https://www.nature.com/articles/s41467-018-07275-x

row.names.remove <- names(subset(total, total <600)) # no samples to remove here
row.names.remove
seqtab.less <- seqtab.trim[!(row.names(seqtab.trim) %in% row.names.remove),]

samdf.rare <- samdf.trim[!(row.names(samdf.trim) %in% row.names.remove), ]
# rarefy to min=600 reads per sample. most reads, except one sample,hace reads higher than 600. 

seqtab.rare <- rrarefy(seqtab.less,sample=600)
pdf("Pacuta_rare_6h.pdf", width=5.5, height=4.5)
rarecurve(seqtab.rare,step=100,label=TRUE)
dev.off()
```


```{r}
#phyloseq object but rarefied
ps.trim.rare <- phyloseq(otu_table(seqtab.rare, taxa_are_rows=FALSE),
                         sample_data(samdf.rare),
                         tax_table(taxa2))
ps.trim.rare #220 taxa and 17 samples

#removing missing taxa - lost after rarefying
ps.trim.rare <- prune_taxa(taxa_sums(ps.trim.rare) > 0, ps.trim.rare)
ps.trim.rare #212 taxa and 17 samples

seqtab.trim.rare <- data.frame(otu_table(ps.trim.rare))

#saving
saveRDS(ps.trim.rare,file="/projectnb/coral/jk_data/project_3_Pacuta/7_Microbiome/adapt_rm_files/degen_rm/clean_reads/phyloseq.trim.rarefied.4k.rds")

write.csv(seqtab.rare, file="/projectnb/coral/jk_data/project_3_Pacuta/7_Microbiome/adapt_rm_files/degen_rm/clean_reads/Pacuta_16s_seqtab.rev.cleanest.rare_4k")

### data files - decontaminated, trimmed, rarefied ####

ps.trim.rare.4k = readRDS("/projectnb/coral/jk_data/project_3_Pacuta/7_Microbiome/adapt_rm_files/degen_rm/clean_reads/phyloseq.trim.rarefied.4k.rds")
seqtab.trim.rare.4k <- data.frame(ps.trim.rare.4k@otu_table)
samdf.trim.rare.4k <- data.frame(ps.trim.rare.4k@sam_data)


load("taxa2.Rdata")

ps.trim.rare <- phyloseq(otu_table(seqtab.trim.rare.4k, taxa_are_rows=FALSE),
                         sample_data(samdf.trim.rare.4k),
                         tax_table(taxa2))
ps.trim.rare # 212 taxa and 17 samples


#### Try Batch Correction ####

library(knitr)
library(xtable) # table
library(sva) # ComBat
library(ggplot2) # PCA sample plot with density
library(gridExtra) # PCA sample plot with density
library(limma) # removeBatchEffect (LIMMA)
library(vegan) # RDA
library(AgiMicroRna) # RLE plot
library(cluster) # silhouette coefficient
library(variancePartition) # variance calculation
library(pvca) # PVCA
library(pheatmap) # heatmap
library(ruv) # RUVIII
library(lmerTest) # lmer
library(bapred) # FAbatch

# data that is cleaned of euk contamination but nothing else
seqtab.cleanest <- data.frame(ps.cleanest@otu_table)
samdf.cleanest <- data.frame(ps.cleanest@sam_data)

# data that is cleaned and rarefied
ps.rare.4k = readRDS("/projectnb/coral/jk_data/project_3_Pacuta/7_Microbiome/adapt_rm_files/degen_rm/clean_reads/phyloseq.trim.rarefied.4k.rds")
samdf.cleanest <- data.frame(ps.rare.4k@sam_data)

# data that is cleaned, trimmed, rarefied
seqtab.cleanest <- data.frame(ps.trim.rare@otu_table)
samdf.cleanest <- data.frame(ps.trim.rare@sam_data)

library(MCMC.OTU)
exists("logratio.transform", where = "package:MCMC.OTU")
library(phyloseq)

## Prefiltering
# ad data
seqtab.index.keep <- which(colSums(seqtab.cleanest)*100/(sum(colSums(seqtab.cleanest))) > 0.01)
seqtab.cleanest.keep <- seqtab.cleanest[, seqtab.index.keep]
dim(seqtab.cleanest.keep)
# [1]  17 193
# Perform CLR (Centered Log-Ratio) transformation
ps_clr <- transform_sample_counts(ps, function(x) log(x + 1))  # Adjust if needed

# Add offset to handle zeros
seqtab.cleanest.keep <- seqtab.cleanest.keep + 1
# Install the package (if not already installed)
library("compositions")
library(MCMC.OTU)
exists("logratio.transform", where = "package:MCMC.OTU")
library(phyloseq)
# Centered log-ratio transformation
seqtab.clr <- logratio.transfo(seqtab.cleanest.keep, logratio = "CLR")
seqtab.clr <- logratio.transfo(seqtab.cleanest.keep, logratio = "CLR")
class(seqtab.clr) <- 'matrix'
```

```{r making phyloseq objects, eval=FALSE,}
# can skip this section once you've run it once and saved output Rdata files
load("taxa2.Rdata")

#cleaned, rarefied files:
ps.rare.4k = readRDS("phyloseq.rarefied.rds")
seqtab.rare.4k <- data.frame(ps.rare.4k@otu_table)
samdf.rare.4k <- data.frame(ps.rare.4k@sam_data)


ps.rare <- phyloseq(otu_table(seqtab.rare.4k, taxa_are_rows=FALSE), 
                    sample_data(samdf.rare.4k), 
                    tax_table(taxa2))
ps.rare 
#213 taxa and 17 samples

save(ps.rare,file="ps.rare.Rdata")

#cleaned, unrarefied files:
ps.clean = readRDS("phyloseq.cleanest.all.rds")
seqtab.clean <- data.frame(ps.clean@otu_table)
samdf.clean <- data.frame(ps.clean@sam_data)

ps.clean <- phyloseq(otu_table(seqtab.clean, taxa_are_rows=FALSE), 
                    sample_data(samdf.clean), 
                    tax_table(taxa2))
ps.clean#231 taxa and 18 samples


save(ps.clean,file="ps.clean.Rdata")

##TRIMMED VERSIONS:
# trimmed and rarefied
ps.trim.rare = readRDS("phyloseq.trim.rarefied.4k.rds")
seqtab.trim.rare <- data.frame(ps.trim.rare@otu_table)
samdf.trim.rare <- data.frame(ps.trim.rare@sam_data)

ps.trim.rare <- phyloseq(otu_table(seqtab.trim.rare, taxa_are_rows=FALSE), 
                    sample_data(samdf.trim.rare), 
                    tax_table(taxa2))
ps.trim.rare# 212 taxa and 17 samples

save(ps.trim.rare,file="ps.trim.rare.Rdata")

# just trimmed
ps.trim = readRDS("phyloseq.cleanest.trim.rds")
seqtab.trim <- data.frame(ps.trim@otu_table)
samdf.trim <- data.frame(ps.trim@sam_data)

ps.trim <- phyloseq(otu_table(seqtab.trim, taxa_are_rows=FALSE), 
                    sample_data(samdf.trim), 
                    tax_table(taxa2))
ps.trim#220 taxa and 18 samples
ps.trim.rare# 212 taxa and 17 samples
ps.cleanest#231 taxa and 18 samples
ps.clean#231 taxa and 18 samples
save(ps.trim,file="ps.trim.Rdata")
```


```{r generate diversity metrics,}
# Diversity
##[Notes from phyloseq author](https://rdrr.io/bioc/phyloseq/man/estimate_richness.html)
#Visualize alpha-diversity - Should be done on raw, untrimmed dataset
#use the cleanest and unrarefied data
df <- data.frame(estimate_richness(ps.cleanest, split=TRUE, measures=c("Shannon","InvSimpson","Observed")))
samdf.clean <- data.frame(ps.cleanest@sam_data)
# Rename values in the Genotype column
samdf.clean <- samdf.clean %>% 
  mutate(Genotype = case_when(
    Genotype == "gen1_" ~ "ECAH",
    Genotype == "gen4_" ~ "ECG",
    Genotype == "gen3_" ~ "ORAPG",
    TRUE ~ Genotype  # Keep the original value if none of the conditions are met
  ))

df$Sample <- rownames(df)
df.div <- merge(df,samdf.clean,by="Sample") #add sample data

#shannon diversity divided by species richness
df.div$even <- df.div$Shannon/(log(df.div$Observed))
str(df.div)
df.div$Treatment = as.factor(df.div$Treatment)
df.div$Algae_type = as.factor(df.div$Algae_type)
df.div$Genotype = as.factor(df.div$Genotype)
#assign colors for algae type
colvec2<- c("ECAH"="brown4","ECG" = "darkgreen", "ORAPG"="darkgreen")
```


```{r shannon lineage,}
## Alpha metrics {.tabset}
#/projectnb/coral/jk_data/project_3_Pacuta/7_Microbiome/adapt_rm_files/degen_rm/clean_reads/
#Looking at everything individually, more of a synthesis below
#Alpha diversity is within-sample diversity. Beta diversity, on the other hand, is a measure of similarity #or dissimilarity between populations/groups. 

# Ensure the 'treatment' column in df.div has the desired order
df.div$Treatment <- factor(df.div$Treatment, levels = c("Control", "Heat"))
### Shannon
#Shannon diversity estimates species diversity, considers number of species (richness) #and their relative #abundance (evenness)
shapes = c("ECAH" = 21, "ECG" = 22, "ORAPG"=24)
colvec<- c("Control" = "#2c7fb8","Heat" = "#de2d26")

library(ggforce)
shan<-ggplot(df.div, aes(x=Treatment, y=Shannon,fill=Treatment,group=Treatment, shape=Genotype))+
  geom_boxplot(outlier.shape=NA, alpha=0.8)+
  scale_fill_manual(values=colvec)+
  scale_color_manual(values=colvec)+
  geom_jitter(size=2.5,width = 0.25, alpha=.9)+
  xlab("Treatment")+
  ylab("Shannon index")+
  theme_classic()+
  theme(axis.text.x = element_text(angle = 45, hjust = 1),legend.position = "none") +
  #ylim(1, 4.5)+
  scale_fill_manual(values = colvec) +
  scale_shape_manual(values = shapes)+facet_wrap(~ Genotype)
shan
```

```{r shannon stats,}
#Shannon
shapiro.test(df.div$Shannon) #
mod<-lm(Shannon~Treatment*Genotype,data=df.div)
anova(mod)
#  Df Sum Sq Mean Sq F value   Pr(>F)   
#Treatment             1 0.0214  0.0214  0.0874 0.772588   
#Algae_type            1 0.3019  0.3019  1.2321 0.288750   
#Genotype              1 2.9166  2.9166 11.9023 0.004805 **
#Treatment:Algae_type  1 0.2072  0.2072  0.8454 0.375964   
##Treatment:Genotype    1 3.7703  3.7703 15.3860 0.002026 **
#Residuals            12 2.9406  0.2450                    
#---
#Signif. codes:  0 '***' 0.001 '**' 0.01 '*' 0.05 '.' 0.1 ' ' 1

#Response: Shannon
#                   Df Sum Sq Mean Sq F value   Pr(>F)   
#Treatment           1 0.0214 0.02141  0.0874 0.772588   
##Genotype            2 3.2185 1.60926  6.5672 0.011844 * 
#Treatment:Genotype  2 3.9775 1.98873  8.1157 0.005898 **
```

```{r simpson,}
### Simpson
#Simpson index considered more of a dominance index, accounts for proportion of species in a sample.
inv<-ggplot(df.div, aes(x=Treatment, y=InvSimpson,fill=Treatment,group=Treatment, shape = Genotype))+
  geom_boxplot(outlier.shape=NA, alpha=0.8)+
  scale_fill_manual(values=colvec)+
  scale_color_manual(values=colvec)+
  geom_jitter(size=2.5,width = 0.25, alpha=.9)+
  xlab("Treatment")+
  ylab("Inverse Simpson Index")+
  theme_classic()+
  theme(axis.text.x = element_text(angle = 45, hjust = 1),legend.position = "none") +
  #ylim(1, 4.5)+
  scale_fill_manual(values = colvec) +
  scale_shape_manual(values = shapes)+facet_wrap(~ Genotype)

inv
```

```{r}
#simpson
shapiro.test(df.div$InvSimpson) #
mod<-lm(InvSimpson~Treatment*Genotype,data=df.div)
anova(mod)
#  Df Sum Sq Mean Sq F value  Pr(>F)  
#Treatment             1  35.17  35.167  1.3114 0.27447  
#Algae_type            1   0.12   0.118  0.0044 0.94824  
#Genotype              1  91.97  91.971  3.4296 0.08879 .
#Treatment:Algae_type  1  22.52  22.515  0.8396 0.37756  
#Treatment:Genotype    1 109.83 109.833  4.0957 0.06584 .
#Residuals            12 321.80  26.816                             
```

```{r,}
### Richness
rich<-ggplot(df.div, aes(x=Treatment, y=Observed,fill=Treatment,group=Treatment, shape = Genotype))+
  geom_boxplot(outlier.shape=NA, alpha=0.8)+
  scale_fill_manual(values=colvec)+
  scale_color_manual(values=colvec)+
  geom_jitter(size=2.5,width = 0.25, alpha=.9)+
  xlab("Treatment")+
  ylab("ASV richness")+
  theme_classic()+
  theme(axis.text.x = element_text(angle = 45, hjust = 1),legend.position = "none") +
  #ylim(1, 4.5)+
  scale_fill_manual(values = colvec) +
  scale_shape_manual(values = shapes)+facet_wrap(~ Algae_type)
rich
```

```{r richness stats,}
#otu richness
shapiro.test(df.div$Observed) #
mod<-lm(Observed~Treatment*Genotype,data=df.div)
anova(mod)
#Response: Observed
#                    Df Sum Sq Mean Sq F value  Pr(>F)  
#Treatment             1 156.06  156.06  2.1296 0.17015  
#Algae_type            1 136.11  136.11  1.8575 0.19795  
#Genotype              1 341.33  341.33  4.6581 0.05187 .
#Treatment:Algae_type  1  28.44   28.44  0.3882 0.54492  
#Treatment:Genotype    1 616.33  616.33  8.4109 0.01332 *
#Residuals            12 879.33   73.28                  
```

```{r evenness,}
even<-ggplot(df.div, aes(x=Treatment, y=even,fill=Treatment,group=Treatment, shape = Genotype))+
  geom_boxplot(outlier.shape=NA, alpha=0.8)+
  scale_fill_manual(values=colvec)+
  scale_color_manual(values=colvec)+
  geom_jitter(size=2.5,width = 0.25, alpha=.9)+
  xlab("Treatment")+
  ylab("Evenness")+
  theme_classic()+
  theme(axis.text.x = element_text(angle = 45, hjust = 1),legend.position = "none") +
  #ylim(1, 4.5)+
  scale_fill_manual(values = colvec) +
  scale_shape_manual(values = shapes)+facet_wrap(~ Algae_type)
even

```
```{r}
#evenness
shapiro.test(df.div$even) #
mod<-lm(even~Treatment*Genotype,data=df.div)
anova(mod)
#Df   Sum Sq  Mean Sq F value   Pr(>F)   
#Treatment             1 0.011632 0.011632  0.9186 0.356746   
#Algae_type            1 0.102665 0.102665  8.1077 0.014693 * 
#Genotype              1 0.184777 0.184777 14.5923 0.002440 **
#Treatment:Algae_type  1 0.016711 0.016711  1.3197 0.273028   
#Treatment:Genotype    1 0.195135 0.195135 15.4103 0.002015 **
#Residuals            12 0.151951 0.012663                          
```


```{r all plot}
plots.lin = ggarrange(shan,inv,even,rich, ncol=2, nrow=2)
#only richness is sig p< 0.05
ggsave(plots.lin, filename = "Pacuta_diversity_plots_genotype_clean_nolabs_MERGE_Facet.pdf", width=8,height=4, units=c("in"), useDingbats=FALSE)
plots.lin
```


```{r packages composition}
library(rlang)
library(stringr)
library(dplyr)
library(stats)
library(ggpubr)
library(vegan)
library(cowplot)
library(tidyverse)
#library(MCMC.OTU)
#install.packages("remotes")
#remotes::install_github("Jtrachsel/funfuns")
library(phyloseq)

```

Read in data

```{r read in data}
load("taxa2.Rdata")

#phyloseq objects
load("ps.clean.Rdata")
load("ps.rare.Rdata")
load("ps.trim.rare.Rdata") # rare otu's trimmed, then rarefied (4k)
ps.clean = readRDS("phyloseq.cleanest.all.rds")

# remove treatment NA's
ps.clean <- subset_samples(ps.clean,(!is.na(Treatment)))
ps.rare <- subset_samples(ps.rare,(!is.na(Treatment)))
ps.trim.rare <- subset_samples(ps.trim.rare,(!is.na(Treatment)))

library(dplyr)
# moving forward with rarefied data, doesn't seem to make a difference in the results whether rare OTUs trimmed or not 

# make the factors of interest actual factors 
sample_data(ps.rare)$Treatment = as.factor(sample_data(ps.rare)$Treatment)
sample_data <- sample_data %>% 
  mutate(Genotype = case_when(
    Genotype == "gen1_" ~ "ECAH",
    Genotype == "gen4_" ~ "ECG",
    Genotype == "gen3_" ~ "ORAPG",
    TRUE ~ Genotype  # Keep the original value if none of the conditions are met
  ))



# set color palettes
shapes = c("ECAH" = 21, "ECG" = 22, "ORAPG"=24)
colvec<- c("Control" = "#2c7fb8","Heat" = "#de2d26")

```

Rename ASVs to be more informative

```{r rename ASVs}
#analyse rarefied first
tax <- as.data.frame(ps.rare@tax_table@.Data)

tax.clean <- data.frame(row.names = row.names(tax),
                        Kingdom = str_replace(tax[,1], "D_0__",""),
                        Phylum = str_replace(tax[,2], "D_1__",""),
                        Class = str_replace(tax[,3], "D_2__",""),
                        Order = str_replace(tax[,4], "D_3__",""),
                        Family = str_replace(tax[,5], "D_4__",""),
                        Genus = str_replace(tax[,6], "D_5__",""),
                        Species = str_replace(tax[,7], "D_6__",""),
                        stringsAsFactors = FALSE)
tax.clean[is.na(tax.clean)] <- ""

for (i in 1:7){ tax.clean[,i] <- as.character(tax.clean[,i])}
tax.clean[is.na(tax.clean)] <- ""
for (i in 1:nrow(tax.clean)){
  if (tax.clean[i,2] == ""){
    kingdom <- paste("Kingdom_", tax.clean[i,1], sep = "")
    tax.clean[i, 2:7] <- kingdom
  } else if (tax.clean[i,3] == ""){
    phylum <- paste("Phylum_", tax.clean[i,2], sep = "")
    tax.clean[i, 3:7] <- phylum
  } else if (tax.clean[i,4] == ""){
    class <- paste("Class_", tax.clean[i,3], sep = "")
    tax.clean[i, 4:7] <- class
  } else if (tax.clean[i,5] == ""){
    order <- paste("Order_", tax.clean[i,4], sep = "")
    tax.clean[i, 5:7] <- order
  } else if (tax.clean[i,6] == ""){
    family <- paste("Family_", tax.clean[i,5], sep = "")
    tax.clean[i, 6:7] <- family
  } else if (tax.clean[i,7] == ""){
    tax.clean$Species[i] <- paste("Genus",tax.clean$Genus[i], sep = "_")
  }
}

tax_table(ps.rare) <- as.matrix(tax.clean)
```


# All ASVs - PCAs

## Rarefied Data

```{r }
# set color palettes
shapes = c("ECAH" = 21, "ECG" = 22, "ORAPG"=24)
colvec<- c("Control" = "#2c7fb8","Heat" = "#de2d26")
ord <- ordinate(ps.rare, "PCoA", "bray")
pcoa.treat=plot_ordination(ps.rare, ord, color = "Treatment", shape = "Genotype") +
  geom_point(size =2, alpha = 0.1) +
  scale_color_manual(values = colvec) +
  stat_ellipse(aes(group = ps.rare@sam_data$Algae_type),lwd = 0.5, size = 0.5) +
  theme_classic()
pcoa.treat
ggsave(pcoa.treat, filename = "Pacuta_pcoas.allasvs.rare__SYMBIONT.pdf", width=3.5, height=2.5, units=c("in"), useDingbats=FALSE)
```
```{r }
# set color palettes
shapes = c("ECAH" = 21, "ECG" = 22, "ORAPG"=24)
colvec<- c("Control" = "#2c7fb8","Heat" = "#de2d26")
ord <- ordinate(ps.rare, "PCoA", "bray")
pcoa.treat=plot_ordination(ps.rare, ord, color = "Treatment", shape = "Genotype") +
  geom_point(size =2, alpha = 0.1) +
  scale_color_manual(values = colvec) +
  stat_ellipse(aes(group = ps.rare@sam_data$Treatment),lwd = 0.5, size = 0.5) +
  theme_classic()
pcoa.treat
ggsave(pcoa.treat, filename = "Pacuta_pcoas.allasvs.rare.pdf", width=3.5, height=2.5, units=c("in"), useDingbats=FALSE)
```

```{r}
tax <- as.data.frame(ps.cleanest@tax_table@.Data)

tax.clean <- data.frame(row.names = row.names(tax),
                        Kingdom = str_replace(tax[,1], "D_0__",""),
                        Phylum = str_replace(tax[,2], "D_1__",""),
                        Class = str_replace(tax[,3], "D_2__",""),
                        Order = str_replace(tax[,4], "D_3__",""),
                        Family = str_replace(tax[,5], "D_4__",""),
                        Genus = str_replace(tax[,6], "D_5__",""),
                        Species = str_replace(tax[,7], "D_6__",""),
                        stringsAsFactors = FALSE)
tax.clean[is.na(tax.clean)] <- ""

for (i in 1:7){ tax.clean[,i] <- as.character(tax.clean[,i])}
tax.clean[is.na(tax.clean)] <- ""
for (i in 1:nrow(tax.clean)){
  if (tax.clean[i,2] == ""){
    kingdom <- paste("Kingdom_", tax.clean[i,1], sep = "")
    tax.clean[i, 2:7] <- kingdom
  } else if (tax.clean[i,3] == ""){
    phylum <- paste("Phylum_", tax.clean[i,2], sep = "")
    tax.clean[i, 3:7] <- phylum
  } else if (tax.clean[i,4] == ""){
    class <- paste("Class_", tax.clean[i,3], sep = "")
    tax.clean[i, 4:7] <- class
  } else if (tax.clean[i,5] == ""){
    order <- paste("Order_", tax.clean[i,4], sep = "")
    tax.clean[i, 5:7] <- order
  } else if (tax.clean[i,6] == ""){
    family <- paste("Family_", tax.clean[i,5], sep = "")
    tax.clean[i, 6:7] <- family
  } else if (tax.clean[i,7] == ""){
    tax.clean$Species[i] <- paste("Genus",tax.clean$Genus[i], sep = "_")
  }
}
```


```{r}
tax_table(ps.cleanest) <- as.matrix(tax.clean)
ord <- ordinate(ps.cleanest, "PCoA", "bray")
pcoa.treat_clean<-plot_ordination(ps.cleanest, ord, color = "Treatment", shape = "Genotype") +
  geom_point(size = 2, alpha = 0.1) +
  scale_color_manual(values = colvec) +
  stat_ellipse(aes(group = ps.cleanest@sam_data$Treatment),lwd = 0.5, size = 0.5) +
  theme_classic()
pcoa.treat_clean

ggsave(pcoa.treat_clean, filename = "Pcatua_pcoas.allasvs.clean_noRARE.pdf", width=3.5, height=2.5, units=c("in"), useDingbats=FALSE)
```
```{r}
tax_table(ps.cleanest) <- as.matrix(tax.clean)
ord <- ordinate(ps.cleanest, "PCoA", "bray")
pcoa.treat_clean<-plot_ordination(ps.cleanest, ord, color = "Treatment", shape = "Genotype") +
  geom_point(size = 2, alpha = 0.1) +
  scale_color_manual(values = colvec) +
  stat_ellipse(aes(group = ps.cleanest@sam_data$Algae_type),lwd = 0.5, size = 0.5) +
  theme_classic()
pcoa.treat_clean

ggsave(pcoa.treat_clean, filename = "Pcatua_pcoas.allasvs.clean_noRARE_SYMBIONT.pdf", width=3.5, height=2.5, units=c("in"), useDingbats=FALSE)
```


### Rarefied Stats

Help on adonis (here)[https://thebiobucket.blogspot.com/2011/04/assumptions-for-permanova-with-adonis.html#more]

```{r stats rarefied data}
seq.rare <- data.frame(otu_table(ps.rare))
seq.clean <- data.frame(otu_table(ps.cleanest))
# make distance matrix with vegdist():
dist.rare <- vegdist(seq.rare,method = "bray")
samdf.rare <- data.frame(sample_data(ps.rare))
row.names(samdf.rare)==row.names(seq.rare)

dist.rare_matrix <- as.matrix(dist.rare)

# Subset the distance matrix for the filtered samples
filtered_dist_matrix <- dist.rare_matrix[rownames(samdf.rare), rownames(samdf.rare)]

# Create a distance object from the filtered distance matrix
filtered_dist <- as.dist(filtered_dist_matrix)
interaction_factor <- interaction(samdf.rare$Treatment, samdf.rare$Genotype)

# Perform beta dispersion analysis based on 'treatment'
bet.all <- betadisper(filtered_dist, interaction_factor)
# Perform ANOVA test
anova_results <- anova(bet.all)

# Print the ANOVA results
print(anova_results)
# Df  Sum Sq  Mean Sq F value Pr(>F)
#Groups     1 0.00071 0.000712  0.0142 0.9068
#Residuals 15 0.75310 0.050206            
#---
#Signif. codes:  0 ‘***’ 0.001 ‘**’ 0.01 ‘*’ 0.05 ‘.’ 0.1 ‘ ’ 1
```

```{r}
# Perform beta dispersion analysis based on 'treatment'
bet.all <- betadisper(filtered_dist, samdf.rare$Algae_type)
# Perform ANOVA test
anova_results <- anova(bet.all)

# Print the ANOVA results
print(anova_results)
#Analysis of Variance Table

#Response: Distances
#          Df  Sum Sq  Mean Sq F value Pr(>F)
#Groups     1 0.00465 0.004651  0.1771 0.6799
#Residuals 15 0.39400 0.026267
```


```{r stats rarefied data}

adonis2(seq.rare ~ Treatment*Genotype, data=samdf.rare, permutations=999)

#   adonis2(formula = seq.rare ~ Treatment * Algae_type * Genotype, data = samdf.rare, permutations = 999)
                     Df SumOfSqs      R2       F Pr(>F)    
#Treatment             1   0.3095 0.07112  2.4654  0.062 .  
#Algae_type            1   0.6903 0.15865  5.4996  0.005 ** 
#Genotype              1   1.6010 0.36793 12.7545  0.001 ***
#Treatment:Algae_type  1   0.2072 0.04762  1.6509  0.149    
#Treatment:Genotype    1   0.1626 0.03737  1.2953  0.241  
#                     Residual             11   1.3807 0.31732                   
#Total                16   4.3513 1.00000                   
#---
#Signif. codes:  0 ‘***’ 0.001 ‘**’ 0.01 ‘*’ 0.05 ‘.’ 0.1 ‘ ’ 1
                     
```
```{r}
adonis2(seq.clean ~ Treatment*Genotype, data=samdf.trim, permutations=999)

#  adonis2(formula = seq.clean ~ Treatment * Algae_type * Genotype, data = samdf.trim, permutations = 999)
#                     Df SumOfSqs      R2      F Pr(>F)    
#Treatment             1   0.3409 0.06352 1.9546  0.062 .  
#Algae_type            1   1.1293 0.21041 6.4744  0.001 ***
#Genotype              1   1.2352 0.23015 7.0816  0.001 ***
#Treatment:Algae_type  1   0.3259 0.06073 1.8687  0.059 .  
#Treatment:Genotype    1   0.2426 0.04520 1.3908  0.178    
#Residual             12   2.0932 0.38999                  
#Total                17   5.3672 1.00000                  
#---
#Signif. codes:  0 ‘***’ 0.001 ‘**’ 0.01 ‘*’ 0.05 ‘.’ 0.1 ‘ ’ 1     
#adonis2(formula = seq.clean ~ Treatment * Genotype, data = samdf.trim, permutations = 999)
#Treatment           1   0.3410 0.06345 1.9460  0.055 .  
#Genotype            2   2.3607 0.43923 6.7352  0.001 ***
#Treatment:Genotype  2   0.5699 0.10603 1.6258  0.060 .
```



```{r stats rarefied data}
seq.clean <- data.frame(otu_table(ps.cleanest))
# make distance matrix with vegdist():
dist.clean <- vegdist(seq.clean,method = "bray")
samdf.clean <- data.frame(sample_data(ps.cleanest))
row.names(samdf.clean)==row.names(seq.clean)

dist.clean_matrix <- as.matrix(dist.clean)

# Subset the distance matrix for the filtered samples
filtered_dist_matrix_cl <- dist.clean_matrix[rownames(samdf.clean), rownames(samdf.clean)]

# Create a distance object from the filtered distance matrix
filtered_dist_cl <- as.dist(filtered_dist_matrix_cl)

# Perform beta dispersion analysis based on 'treatment'
bet.all_cl <- betadisper(filtered_dist_cl, samdf.clean$Treatment)
# Perform ANOVA test
anova_results <- anova(bet.all_cl)

# Print the ANOVA results
print(anova_results)
#Analysis of Variance Table

#Df   Sum Sq   Mean Sq F value Pr(>F)
#Groups     1 0.008574 0.0085739  0.6453 0.4336
#Residuals 16 0.212582 0.0132864
```

```{r}
# Perform beta dispersion analysis based on 'treatment'
bet.all_cl <- betadisper(filtered_dist_cl, samdf.clean$Algae_type)
# Perform ANOVA test
anova_results <- anova(bet.all_cl)

# Print the ANOVA results
print(anova_results)
#Response: Distances
#          Df   Sum Sq   Mean Sq F value Pr(>F)
#Groups     1 0.000465 0.0004653  0.0343 0.8554
#Residuals 16 0.217162 0.0135726     
```


```{r stats rarefied data}
adonis2(formula = seq.clean ~ Treatment*Genotype, data = samdf.clean, permutations = 999)
#Df SumOfSqs      R2      F Pr(>F)    
##Treatment             1   0.3410 0.06345 1.9460  0.058 .  
#Algae_type            1   1.1264 0.20958 6.4273  0.001 ***
#Genotype              1   1.2343 0.22965 7.0430  0.001 ***
#Treatment:Algae_type  1   0.3270 0.06083 1.8657  0.057 .  
#Treatment:Genotype    1   0.2429 0.04519 1.3860  0.194    
#Residual             12   2.1030 0.39129                  
#Total                17   5.3747 1.00000  
```

```{r}
# Remove taxa where any taxonomic information is NA
ps.trim <- prune_taxa(!apply(tax_table(ps.trim), 1, function(x) any(is.na(x))), ps.trim)

phylo_trim_rel <- prune_samples(sample_sums(ps.trim) > 0, ps.trim)
phylo_trim_rel <- transform_sample_counts(phylo_trim_rel, function(x) x / sum(x))
# Assuming ps.trim is your phyloseq object
pac_ph<-plot_bar(phylo_trim_rel, x = "Sample", fill="Genus") +
   geom_bar(stat = "identity", color = NA) +
  theme_bw() +
  theme(panel.grid = element_blank(), legend.position = "bottom") +
  scale_fill_manual(values = rev(colorRampPalette(brewer.pal(8, "Paired"))(75))) +
  scale_colour_manual(values = rev(colorRampPalette(brewer.pal(8, "Paired"))(75))) +
  ylab("Relative abundance") +
  theme(axis.text.x = element_text(angle = 45, hjust = 1))+facet_wrap(~Genotype, scales = "free_x") 
pac_ph
ggsave(pac_ph, filename = "Pacuta_All_phyla.trim_genus.pdf", width=8, height=7, units=c("in"), useDingbats=FALSE)

```
```{r}
# Remove taxa where any taxonomic information is NA
ps.trim <- prune_taxa(!apply(tax_table(ps.trim), 1, function(x) any(is.na(x))), ps.trim)

phylo_trim_rel <- prune_samples(sample_sums(ps.trim) > 0, ps.trim)
phylo_trim_rel <- transform_sample_counts(phylo_trim_rel, function(x) x / sum(x))
# Assuming ps.trim is your phyloseq object
pac_ph<-plot_bar(phylo_trim_rel, x = "Sample", fill="Genus") +
   geom_bar(stat = "identity", color = NA) +
  theme_bw() +
  theme(panel.grid = element_blank(), legend.position = "bottom") +
  scale_fill_manual(values = rev(colorRampPalette(brewer.pal(8, "Paired"))(75))) +
  scale_colour_manual(values = rev(colorRampPalette(brewer.pal(8, "Paired"))(75))) +
  ylab("Relative abundance") +
  theme(axis.text.x = element_text(angle = 45, hjust = 1))+facet_wrap(~Genotype, scales = "free_x") 
pac_ph
ggsave(pac_ph, filename = "Pacuta_All_phyla.trim_genus.pdf", width=8, height=7, units=c("in"), useDingbats=FALSE)

```


```{r}
# Assuming ps.trim is your phyloseq object
pac_ph_fam<-plot_bar(phylo_trim_rel, x = "Sample", fill="Family") +
   geom_bar(stat = "identity", color = NA) +
  theme_bw() +
  theme(panel.grid = element_blank(), legend.position = "bottom") +
  scale_fill_manual(values = rev(colorRampPalette(brewer.pal(8, "Paired"))(56))) +
  scale_colour_manual(values = rev(colorRampPalette(brewer.pal(8, "Paired"))(56))) +
  ylab("Relative abundance") +
  theme(axis.text.x = element_text(angle = 45, hjust = 1))+facet_wrap(~Genotype, scales = "free_x") 
pac_ph_fam
ggsave(pac_ph_fam, filename = "Pacuta_All_phyla.trim_Family.pdf", width=8, height=10, units=c("in"), useDingbats=FALSE)

```
```{r}
# Assuming ps.trim is your phyloseq object
pac_ph<-plot_bar(phylo_trim_rel, x = "Sample", fill="Phylum") +
   geom_bar(stat = "identity", color = NA) +
  theme_bw() +
  theme(panel.grid = element_blank(), legend.position = "bottom") +
  scale_fill_manual(values = rev(colorRampPalette(brewer.pal(8, "Paired"))(9))) +
  scale_colour_manual(values = rev(colorRampPalette(brewer.pal(8, "Paired"))(9))) +
  ylab("Relative abundance") +
  theme(axis.text.x = element_text(angle = 45, hjust = 1))+facet_wrap(~Genotype, scales = "free_x") 
pac_ph
ggsave(pac_ph, filename = "Pacuta_All_phyla.trim_Phylum.pdf", width=8, height=4, units=c("in"), useDingbats=FALSE)

```


```{r}
## Selecting just Phylum Bacteroidota which is the dominant phyla
phylo_proteo <- subset_taxa(ps.trim, Phylum %in% "Bacteroidota")
phylo_proteo_rel <- transform_sample_counts(phylo_proteo, function(x) x / sum(x))

Proteo_ai<-plot_bar(phylo_proteo_rel, x = "Sample", fill = "Order")+
 geom_bar(stat = "identity", color = NA) +
  theme_bw() +
  theme(panel.grid = element_blank(), legend.position = "bottom", legend.key.size = unit(0.3, 'cm')) +
  scale_fill_manual(values = rev(colorRampPalette(brewer.pal(8, "Paired"))(15))) +
  scale_colour_manual(values = rev(colorRampPalette(brewer.pal(8, "Paired"))(15))) +
  ylab("Relative abundance") +
  ggtitle("Bacteroidota ASVs (relative abundance)")+
  theme(axis.text.x = element_text(angle = 45, hjust = 1))+facet_wrap(~Genotype, scales = "free_x") 
Proteo_ai
ggsave(Proteo_ai, filename = "Pacuta_Bacteroidota.trim.pdf", width=3.2, height=4.5, units=c("in"), useDingbats=FALSE)
```

```{r}
## Selecting just Phylum Proteobacteria 
phylo_proteo.rare <- subset_taxa(ps.rare, Phylum %in% "Proteobacteria")
phylo_proteo_rel.rare <- transform_sample_counts(phylo_proteo.rare, function(x) x / sum(x))

Proteo.rare<-plot_bar(phylo_proteo_rel.rare, x = "Sample", fill = "Order")+
 geom_bar(stat = "identity", color = NA) +
  theme_bw() +
  theme(panel.grid = element_blank(), legend.position = "bottom", legend.key.size = unit(0.3, 'cm')) +
  scale_fill_manual(values = rev(colorRampPalette(brewer.pal(8, "Paired"))(25))) +
  scale_colour_manual(values = rev(colorRampPalette(brewer.pal(8, "Paired"))(25))) +
  ylab("Relative abundance") +
  ggtitle("Proteobacteria  ASVs (relative abundance)")+
  theme(axis.text.x = element_text(angle = 45, hjust = 1))+facet_wrap(~Genotype, scales = "free_x") 
Proteo.rare
ggsave(Proteo.rare, filename = "Pacuta_Proteo.trim.rare.pdf", width=3.2, height=4.5, units=c("in"), useDingbats=FALSE)
```

```{r}
# Combine the data frames
# Subset both datasets for Proteobacteria
phylo_proteo_aip <- subset_taxa(ps.rare, Phylum == "Proteobacteria")
phylo_proteo_ste <- subset_taxa(ps.rare_ste, Phylum == "Proteobacteria")

# Calculate relative abundance
rel_proteo_aip <- transform_sample_counts(proteo_aip, function(x) x / sum(x))
rel_proteo_ste <- transform_sample_counts(proteo_ste, function(x) x / sum(x))

# Convert to data frames
df_aip <- psmelt(rel_proteo_aip)
df_ste <- psmelt(rel_proteo_ste)

# Identify common and unique taxa based on Order
common_orders <- intersect(df_aip$Order, df_ste$Order)
unique_aip_orders <- setdiff(df_aip$Order, df_ste$Order)
unique_ste_orders <- setdiff(df_ste$Order, df_aip$Order)

# Adjust taxa names, marking unique taxa as 'NA' for visualization purposes
df_aip$Order <- ifelse(df_aip$Order %in% unique_aip_orders, NA, df_aip$Order)
df_ste$Order <- ifelse(df_ste$Order %in% unique_ste_orders, NA, df_ste$Order)

# Include NA in the common list for color mapping
all_orders <- union(common_orders, c(NA))

# Prepare color palette
color_palette <- colorRampPalette(brewer.pal(min(8, length(all_orders)), "Paired"))(length(all_orders) - 1)
colors <- setNames(c(color_palette, "grey"), all_orders)

# Plot for Aiptasia
plot_aip <- ggplot(df_aip, aes(x = Sample, y = Abundance, fill = Order)) +
  geom_bar(stat = "identity", position = "stack") +
  scale_fill_manual(values = colors, na.value = "grey") +
  theme_minimal() +
 theme(panel.grid = element_blank(), legend.position = "none", legend.key.size = unit(0.1, 'cm')) +
  labs(title = "Proteobacteria ASVs", y = "Relative Abundance")

# Plot for Stella
plot_ste <- ggplot(df_ste, aes(x = Sample, y = Abundance, fill = Order)) +
  geom_bar(stat = "identity", position = "stack") +
  scale_fill_manual(values = colors, na.value = "grey") +
  theme_minimal() +
  theme(panel.grid = element_blank(), legend.position = "none", legend.key.size = unit(0.1, 'cm')) +
  labs(title = "Proteobacteria ASVs", y = "Relative Abundance")

# Display plots side by side
combine<-grid.arrange(plot_ste, plot_aip, nrow = 1)
ggsave(combine, filename = "Proteo.relative_abundance_rare_v1.pdf", width=5, height=4.5, units=c("in"), useDingbats=FALSE)

```


```{r}
phyloseq.cleanest <- "phyloseq.cleanest.all.rds"  # Replace with your actual
# Now use the variable to load the RDS file
acuta_data <- readRDS(phyloseq.cleanest)
```

```{r}
#bin algal type Cladocopium and check pcoa structure
#Subset to only Algal_type “C”
pclean_C <- subset_samples(acuta_data, Algae_type == "C")
colvec<- c("Control" = "grey40","Heat" = "coral1")
# Compute PCoA on your C-only object
ord_C <- ordinate(pclean_C, method = "PCoA", distance = "bray")

# Plot
pcoa.C <- plot_ordination(pclean_C, ord_C, color = "Treatment") +
  geom_point(size = 2, alpha = 0.8) +
  scale_color_manual(values = colvec) +
  stat_ellipse(aes(group = Treatment), 
               lwd = 0.5, linetype = "dashed", show.legend = FALSE) +
  theme_classic() +
  labs(title = "PCoA of C-hosting corals",
       x = paste0("PCoA1 (", round(ord_C$values[1,2] * 100, 1), "%)"),
       y = paste0("PCoA2 (", round(ord_C$values[2,2] * 100, 1), "%)"))

pcoa.C

```
```{r}
library(dplyr)
library(ggplot2)

#retrieve plotting data behind plot_ordination()
base_plot <- plot_ordination(pclean_C, ord_C, color = "Treatment")
ord_df    <- base_plot$data

# compute one hull per Treatment
hulls <- ord_df %>%
  group_by(Treatment) %>%
  slice(chull(Axis.1, Axis.2))

# Draw points + hulls
pcoa.Cv1=ggplot(ord_df, aes(x = Axis.1, y = Axis.2, color = Treatment, fill = Treatment)) +
  geom_point(size = 2, alpha = 0.8) +
  scale_color_manual(values = colvec) +
  scale_fill_manual(values = colvec) +
  geom_polygon(data = hulls,
               aes(group = Treatment),
               alpha = 0.2,
               show.legend = FALSE) +
  theme_classic() +
  labs(title = "PCoA of C-hosting corals",
       x = paste0("PCoA1 (", round(ord_C$values[1,2] * 100, 1), "%)"),
       y = paste0("PCoA2 (", round(ord_C$values[2,2] * 100, 1), "%)"))
pcoa.Cv1
```

```{r}
seq.clean_C <- data.frame(otu_table(pclean_C))
# make distance matrix with vegdist():
dist.clean_C <- vegdist(seq.clean_C,method = "bray")
samdf.clean_C <- data.frame(sample_data(pclean_C))
row.names(samdf.clean_C)==row.names(seq.clean_C)

dist.clean_matrix_C <- as.matrix(dist.clean_C)

# Subset the distance matrix for the filtered samples
filtered_dist_matrix_cl_C <- dist.clean_matrix_C[rownames(samdf.clean_C), rownames(samdf.clean_C)]

# Create a distance object from the filtered distance matrix
filtered_dist_cl_C <- as.dist(filtered_dist_matrix_cl_C)

# Perform beta dispersion analysis based on 'treatment'
bet.all_cl_C <- betadisper(filtered_dist_cl_C, samdf.clean_C$Treatment)
# Perform ANOVA test
anova_results_C <- anova(bet.all_cl_C)

# Print the ANOVA results
print(anova_results_C)
#Response: Distances
 #         Df   Sum Sq  Mean Sq F value Pr(>F)
#Groups     1 0.009902 0.009902  0.1822 0.6915
#Residuals  4 0.217381 0.054345
```

```{r stats rarefied data}
adonis2(formula = seq.clean_C ~ Treatment, data = samdf.clean_C, permutations = 999)
#adonis2(formula = seq.clean_C ~ Treatment, data = samdf.clean_C, permutations = 999)
#          Df SumOfSqs      R2      F Pr(>F)
#Treatment  1  0.33671 0.22625 1.1696    0.4
#Residual   4  1.15152 0.77375              
#Total      5  1.48823 1.00000
```

```{r}
#bin algal type Cladocopium and check pcoa structure
#Subset to only Algal_type Durusdinium
pclean_D <- subset_samples(acuta_data, Algae_type == "D")
colvec<- c("Control" = "grey40","Heat" = "coral1")
# Compute PCoA on your C-only object
ord_D <- ordinate(pclean_D, method = "PCoA", distance = "bray")

# Plot
pcoa.D <- plot_ordination(pclean_D, ord_D, color = "Treatment", shape = "Genotype") +
  geom_point(size = 2, alpha = 0.8) +
  scale_color_manual(values = colvec) +
  stat_ellipse(aes(group = Treatment), 
               lwd = 0.5, linetype = "dashed", show.legend = FALSE) +
  theme_classic() +
  labs(title = "PCoA of D-hosting corals",
       x = paste0("PCoA1 (", round(ord_C$values[1,2] * 100, 1), "%)"),
       y = paste0("PCoA2 (", round(ord_C$values[2,2] * 100, 1), "%)"))

pcoa.D
```
```{r}
seq.clean_D <- data.frame(otu_table(pclean_D))
# make distance matrix with vegdist():
dist.clean_D <- vegdist(seq.clean_D,method = "bray")
samdf.clean_D <- data.frame(sample_data(pclean_D))
row.names(samdf.clean_D)==row.names(seq.clean_D)

dist.clean_matrix_D <- as.matrix(dist.clean_D)

# Subset the distance matrix for the filtered samples
filtered_dist_matrix_cl_D <- dist.clean_matrix_D[rownames(samdf.clean_D), rownames(samdf.clean_D)]

# Create a distance object from the filtered distance matrix
filtered_dist_cl_D <- as.dist(filtered_dist_matrix_cl_D)

# Perform beta dispersion analysis based on 'treatment'
bet.all_cl_D <- betadisper(filtered_dist_cl_D, samdf.clean_D$Treatment)
# Perform ANOVA test
anova_results_D <- anova(bet.all_cl_D)

# Print the ANOVA results
print(anova_results_D)
#Response: Distances
#          Df   Sum Sq  Mean Sq F value Pr(>F)
#Groups     1 0.027691 0.027690   2.183 0.1703
#Residuals 10 0.126845 0.012684
```

```{r stats rarefied data}
adonis2(formula = seq.clean_D ~ Treatment*Genotype, data = samdf.clean_D, permutations = 999)
#adonis2(formula = seq.clean_D ~ Treatment * Genotype, data = samdf.clean_D, permutations = 999)
#                   Df SumOfSqs      R2       F Pr(>F)    
#Treatment           1  0.33130 0.12003  2.7854  0.052 .  
#Genotype            1  1.23431 0.44721 10.3776  0.001 ***
#Treatment:Genotype  1  0.24289 0.08800  2.0422  0.097 .  
#Residual            8  0.95152 0.34475                   
##Total              11  2.76002 1.00000                   
#---
#Signif. codes:  0 ‘***’ 0.001 ‘**’ 0.01 ‘*’ 0.05 ‘.’ 0.1 ‘ ’ 1
```


```{r }
combined_pcoa <- ggarrange(pcoa.Cv1,pcoa.D, 
                           ncol = 2, nrow = 1, # Define the layout: 2 columns, 2 rows
                           labels = c("A", "B"), # Label each plot
                           common.legend = TRUE, legend = "bottom") # Common legend at the bottom

# Print the combined plot
print(combined_pcoa)
ggsave(combined_pcoa, filename = "Pacuta_pcoas.DandC.pdf", width=5.5, height=2.5, units=c("in"), useDingbats=FALSE)
```


```{r}
# Subset both datasets for Proteobacteria
# You need to filter more specifically, assuming 'Phylum' level is used here for demonstration
all_acuta <- subset_taxa(acuta_data, Kingdom == "Bacteria")

# Calculate relative abundance
rel_all_acuta <- transform_sample_counts(all_acuta, function(x) x / sum(x))

# Convert to data frames
df_acuta <- psmelt(rel_all_acuta)

# Add a source column to differentiate data
df_acuta$source <- 'acuta'

library(ggplot2)
library(RColorBrewer)

# Assign colors
color_proteo <- "red"  # Color for Proteobacteria
other_colors <- brewer.pal(8, "Set2")  # Adjust as necessary

# Function to assign colors, ensuring other taxa are not just grey
assign_colors <- function(df) {
  unique_phyla <- unique(df$Phylum)
  # Generate enough colors for all phyla, repeating palette if necessary
  color_palette <- brewer.pal(8, "Paired")
  num_repeats <- ceiling(length(unique_phyla) / length(color_palette))
  extended_palette <- rep(color_palette, num_repeats)

  # Assign colors to each phylum
  colors <- setNames(extended_palette[1:length(unique_phyla)], unique_phyla)
  colors["Proteobacteria"] <- "darkgreen"  # Ensuring Proteobacteria are always red
  return(colors)
}

```

```{r}
# Subset to only include Phylum Gammaproteobacteria
gamma_acuta <- subset_taxa(acuta_data, Class == "Gammaproteobacteria")

# Transform the data to relative abundances if not already
gamma_acuta_rel <- transform_sample_counts(gamma_acuta, function(x) x / sum(x))

# Aggregate data at the Order level
gamma_orders_st <- tax_glom(gamma_acuta_rel, "Family")

# Extract abundance data
order_abundances_gamma <- psmelt(gamma_orders_st)
#control order of x axis
order_abundances_gamma$treatment <- factor(order_abundances_gamma$Treatment,
                                        levels = c("Control", "Heat"))  


colvec <- c("Heat" = "coral1", "Control" = "grey")
shapes = c("gen1_" = 21, "gen3_" = 22, "gen4_"=24)

   
# Create the boxplot with facets
  
gamma_box<-ggplot(data=order_abundances_gamma, aes(x = Treatment, y = Abundance,fill = Treatment, group = Treatment, shape=Genotype)) +
  geom_boxplot(outlier.shape = NA, alpha = 0.8) +
   geom_jitter(aes(color = Treatment), size = 2.5, width = 0.15, alpha = 0.6) +
  labs(x = "", y = "Abundance\n") +
  scale_shape_manual(values = shapes)+
  scale_color_manual(values = colvec) +
  scale_fill_manual(values = colvec) +
  facet_wrap(~Family, scales = "free")
gamma_box

ggsave(gamma_box, filename = "Proteo_gamma_box.pdf", width=8, height=8, units=c("in"), useDingbats=FALSE)
```


```{r}
# Calculate the total abundance and Bacteroidia abundance for df_aip
total_abundance_acuta <- aggregate(Abundance ~ Sample, data = df_acuta, FUN = sum)
bacte_abundance_acuta <- aggregate(Abundance ~ Sample, data = df_acuta[df_acuta$Class == "Bacteroidia", ], FUN = sum)

# Merge and calculate the percentage for Bacteroidia
abundance_data_acuta <- merge(total_abundance_acuta, bacte_abundance_acuta, by = "Sample", suffixes = c(".Total", ".Bacte"))
abundance_data_acuta$Bacte_Percentage <- (abundance_data_acuta$Abundance.Bacte / abundance_data_acuta$Abundance.Total) * 100

# Replace NA with 0 if there are samples with no Bacteroidia
abundance_data_acuta$Bacte_Percentage[is.na(abundance_data_acuta$Bacte_Percentage)] <- 0

```

```{r}
library(dplyr)
library(stringr)

# Add columns to the matすにさTreatment,Genotype, Algae_type column to 

abundance_data_acuta <- abundance_data_acuta %>%
  mutate(Algae_type = ifelse(grepl("^ECAH", Sample), "Cladocopium",
                        ifelse(grepl("^(ECG|ORAPG)", Sample), "Durusdinium", NA)))


abundance_data_acuta <- abundance_data_acuta %>%
  mutate(Genotype = str_extract(Sample, "^(ECAH|ECG|ORAPG)"))
abundance_data_acuta <- abundance_data_acuta %>%
  mutate(Treatment = case_when(
    str_detect(Sample, "(1|2|3)$") ~ "Control",
    str_detect(Sample, "(4|5|6)$") ~ "Heat",
    TRUE                           ~ NA_character_
  ))

```

```{r}
# #apply lm anova
abundance_data_acuta_bacte <- lm(Bacte_Percentage~ Treatment*Algae_type*Genotype, data = abundance_data_acuta)
# Perform ANOVA on results
anova(abundance_data_acuta_bacte)
#Analysis of Variance Table
#Response: Proteo_Percentage
#Response: Bacte_Percentage
#                     Df Sum Sq Mean Sq F value   Pr(>F)   
#Treatment             1 1358.8  1358.8  5.9971 0.030655 * 
#Algae_type            1  282.1   282.1  1.2449 0.286379   
#Genotype              1 3288.3  3288.3 14.5128 0.002486 **
#Treatment:Algae_type  1  415.8   415.8  1.8353 0.200458   
#Treatment:Genotype    1  386.2   386.2  1.7043 0.216200   
#Residuals            12 2718.9   226.6                    
#---
#Signif. codes:  0 ‘***’ 0.001 ‘**’ 0.01 ‘*’ 0.05 ‘.’ 0.1 ‘ ’ 1

kruskal.test(Bacte_Percentage ~ Algae_type, data = abundance_data_acuta)
```
```{r}
library(emmeans)

#Get estimated marginal means
emm_b <- emmeans(abundance_data_acuta_bacte,
               ~ Treatment*Algae_type*Genotype)

# Pairwise contrasts across one factor (e.g. Treatment),
#     holding the others at their combinations:
pairs(emm_b, by = c("Algae_type","Genotype"), adjust = "tukey")

#Algae_type = Cladocopium, Genotype = ECAH:
# contrast       estimate   SE df t.ratio p.value
# Control - Heat     3.78 12.3 12   0.308  0.9494

#Algae_type = Durusdinium, Genotype = ECG:
# contrast       estimate   SE df t.ratio p.value
# Control - Heat    12.83 12.3 12   1.044  0.5649

#Algae_type = Durusdinium, Genotype = ORAPG:
# contrast       estimate   SE df t.ratio p.value
# Control - Heat    35.52 12.3 12   2.890  0.0337

```

```{r}
# Calculate the total abundance and Gammaproteobacteria abundance for df_aip
gamma_abundance_acuta <- aggregate(Abundance ~ Sample, data = df_acuta[df_acuta$Class == "Gammaproteobacteria", ], FUN = sum)

# Merge and calculate the percentage for Gammaproteobacteria
abundance_data_acuta <- merge(total_abundance_acuta, gamma_abundance_acuta, by = "Sample", suffixes = c(".Total", ".Gamma"))
abundance_data_acuta$Gamma_Percentage <- (abundance_data_acuta$Abundance.Gamma / abundance_data_acuta$Abundance.Total) * 100

# Replace NA with 0 if there are samples with no Gammaproteobacteria
abundance_data_acuta$Gamma_Percentage[is.na(abundance_data_acuta$Gamma_Percentage)] <- 0
```

```{r}
# Add columns to the matすにさTreatment,Genotype, Algae_type column to 

abundance_data_acuta <- abundance_data_acuta %>%
  mutate(Algae_type = ifelse(grepl("^ECAH", Sample), "Cladocopium",
                        ifelse(grepl("^(ECG|ORAPG)", Sample), "Durusdinium", NA)))



abundance_data_acuta <- abundance_data_acuta %>%
  mutate(Genotype = str_extract(Sample, "^(ECAH|ECG|ORAPG)"))
abundance_data_acuta <- abundance_data_acuta %>%
  mutate(Treatment = case_when(
    str_detect(Sample, "(1|2|3)$") ~ "Control",
    str_detect(Sample, "(4|5|6)$") ~ "Heat",
    TRUE                           ~ NA_character_
  ))

```

```{r}

   
colvec <- c("Control" = "gray40", "Heat" = "coral1")
shapes = c("ECAH" = 21, "ECG" = 22, "ORAPG"=24)
acuta_bacte<-ggplot(abundance_data_acuta, aes(x = Treatment, y = Bacte_Percentage, fill = Treatment, group = Treatment, shape=Genotype)) +
  geom_boxplot(outlier.shape = NA, alpha = 0.8) +
  scale_fill_manual(values = colvec) +
  scale_shape_manual(values = shapes)+
  scale_color_manual(values = colvec) +
  geom_jitter(aes(color = Treatment), size = 2.5, width = 0.15, alpha = 0.6) +
  labs(x = "Treatment", y = "Percentage of Bacteroidia") +  # Corrected to use 'labs' for both labels
  theme_classic() +
  theme(axis.text.x = element_text(angle = 45, hjust = 1), legend.position = "none") +
  ylim(-1, 75)+facet_wrap(~Genotype)  # Set y-axis limits from 1 to 100
acuta_bacte

```


```{r}
# #apply lm anova
abundance_data_acuta_gamma<- lm(Gamma_Percentage~ Treatment*Algae_type*Genotype, data = abundance_data_acuta)
# Perform ANOVA on results
anova(abundance_data_acuta_gamma)
#Analysis of Variance Table
#Response: Proteo_Percentage
#Response: Bacte_Percentage
#Response: Gamma_Percentage
#                     Df  Sum Sq Mean Sq F value  Pr(>F)  
#Treatment             1   40.90   40.90  0.2616 0.61830  
#Algae_type            1  283.07  283.07  1.8108 0.20329  
#Genotype              1  687.60  687.60  4.3986 0.05782 .
#Treatment:Algae_type  1   32.63   32.63  0.2088 0.65591  
#Treatment:Genotype    1  193.46  193.46  1.2376 0.28773  
#Residuals            12 1875.88  156.32                  
#---
#Signif. codes:  0 ‘***’ 0.001 ‘**’ 0.01 ‘*’ 0.05 ‘.’ 0.1 ‘ ’ 1

kruskal.test(Bacte_Percentage ~ Algae_type, data = abundance_data_acuta)
```

```{r}
library(emmeans)

#Get estimated marginal means
emm_g <- emmeans(abundance_data_acuta_gamma,
               ~ Treatment*Algae_type*Genotype)

# Pairwise contrasts across one factor (e.g. Treatment),
#     holding the others at their combinations:
pairs(emm_g, by = c("Algae_type","Genotype"), adjust = "tukey")

#Algae_type = Cladocopium, Genotype = ECAH:
# contrast       estimate   SE df t.ratio p.value
# Control - Heat     6.82 10.2 12   0.668  0.7858

#Algae_type = Durusdinium, Genotype = ECG:
# contrast       estimate   SE df t.ratio p.value
# Control - Heat     9.14 10.2 12   0.895  0.6533

#Algae_type = Durusdinium, Genotype = ORAPG:
# contrast       estimate   SE df t.ratio p.value
# Control - Heat    -6.92 10.2 12  -0.678  0.7805
```



```{r}
acuta_gamma<-ggplot(abundance_data_acuta, aes(x = Treatment, y = Gamma_Percentage, fill = Treatment, group = Treatment, shape=Genotype)) +
  geom_boxplot(outlier.shape = NA, alpha = 0.8) +
  scale_fill_manual(values = colvec) +
  scale_shape_manual(values = shapes)+
  scale_color_manual(values = colvec) +
  geom_jitter(aes(color = Treatment), size = 2.5, width = 0.15, alpha = 0.6) +
  labs(x = "Treatment", y = "Percentage of Gammaproteobacteria") +  # Corrected to use 'labs' for both labels
  theme_classic() +
  theme(axis.text.x = element_text(angle = 45, hjust = 1), legend.position = "none") +
  ylim(-1, 55)+facet_wrap(~Genotype)  # Set y-axis limits from 1 to 100
acuta_gamma
```

```{r}
# Calculate the total abundance and Alphaproteobacteria abundance for df_aip
alpha_abundance_acuta <- aggregate(Abundance ~ Sample, data = df_acuta[df_acuta$Class == "Alphaproteobacteria", ], FUN = sum)

# Merge and calculate the percentage for Alphaproteobacteria
abundance_data_acuta <- merge(total_abundance_acuta, alpha_abundance_acuta, by = "Sample", suffixes = c(".Total", ".Alpha"))
abundance_data_acuta$Alpha_Percentage <- (abundance_data_acuta$Abundance.Alpha / abundance_data_acuta$Abundance.Total) * 100

# Replace NA with 0 if there are samples with no Alphaproteobacteria
abundance_data_acuta$Alpha_Percentage[is.na(abundance_data_acuta$Alpha_Percentage)] <- 0
```

```{r}

# Add columns to the matすにさTreatment,Genotype, Algae_type column to 

abundance_data_acuta <- abundance_data_acuta %>%
  mutate(Algae_type = ifelse(grepl("^ECAH", Sample), "Cladocopium",
                        ifelse(grepl("^(ECG|ORAPG)", Sample), "Durusdinium", NA)))


abundance_data_acuta <- abundance_data_acuta %>%
  mutate(Genotype = str_extract(Sample, "^(ECAH|ECG|ORAPG)"))
abundance_data_acuta <- abundance_data_acuta %>%
  mutate(Treatment = case_when(
    str_detect(Sample, "(1|2|3)$") ~ "Control",
    str_detect(Sample, "(4|5|6)$") ~ "Heat",
    TRUE                           ~ NA_character_
  ))

```



```{r}
# #apply lm anova
abundance_data_acuta_alpha<- lm(Alpha_Percentage~ Treatment*Algae_type*Genotype, data = abundance_data_acuta)
# Perform ANOVA on results
anova(abundance_data_acuta_alpha)
#Response: Alpha_Percentage
#                    Df Sum Sq Mean Sq F value    Pr(>F)    
#Treatment             1   0.99    0.99  0.1991 0.6633906    
#Algae_type            1   3.53    3.53  0.7101 0.4158899    
##Genotype              1 434.27  434.27 87.3371 7.405e-07 ***
#Treatment:Algae_type  1 153.60  153.60 30.8915 0.0001242 ***
#Treatment:Genotype    1 229.69  229.69 46.1931 1.912e-05 ***
#Residuals            12  59.67    4.97                      
```

```{r}
#Get estimated marginal means
emm_a <- emmeans(abundance_data_acuta_alpha,
               ~ Treatment*Algae_type*Genotype)

# Pairwise contrasts across one factor (e.g. Treatment),
#     holding the others at their combinations:
pairs(emm_a, by = c("Algae_type","Genotype"), adjust = "tukey")

#Algae_type = Cladocopium, Genotype = ECAH:
# contrast       estimate   SE df t.ratio p.value
# Control - Heat     7.79 1.82 12   4.280  0.0028

##Algae_type = Durusdinium, Genotype = ECG:
# contrast       estimate   SE df t.ratio p.value
# Control - Heat     4.15 1.82 12   2.279  0.0977

#Algae_type = Durusdinium, Genotype = ORAPG:
# contrast       estimate   SE df t.ratio p.value
# Control - Heat   -13.35 1.82 12  -7.333  <.0001
#Signif. codes:  0 ‘***’ 0.001 ‘**’ 0.01 ‘*’ 0.05 ‘.’ 0.1 ‘ ’ 1
```


```{r}
colvec <- c("Control" = "gray40", "Heat" = "coral1")
shapes = c("ECAH" = 21, "ECG" = 22, "ORAPG"=24)
acuta_alpha<-ggplot(abundance_data_acuta, aes(x = Treatment, y = Alpha_Percentage, fill = Treatment, group = Treatment, shape=Genotype)) +
  geom_boxplot(outlier.shape = NA, alpha = 0.8) +
  scale_fill_manual(values = colvec) +
  scale_shape_manual(values = shapes)+
  scale_color_manual(values = colvec) +
  geom_jitter(aes(color = Treatment), size = 2.5, width = 0.15, alpha = 0.6) +
  labs(x = "Treatment", y = "Percentage of Bacteroidia") +  # Corrected to use 'labs' for both labels
  theme_classic() +
  theme(axis.text.x = element_text(angle = 45, hjust = 1), legend.position = "none") +
  ylim(-1, 21)+facet_wrap(~Genotype)  # Set y-axis limits from 1 to 100
acuta_alpha
```

```{r}
library(gridExtra)
# Display plots side by side
combine<-grid.arrange(acuta_bacte, acuta_gamma,acuta_alpha, nrow = 1)
ggsave(combine, filename = "Percen_Class.pdf", width=10, height=2.5, units=c("in"), useDingbats=FALSE)

```

